### 2024-01-04 paper/main_sections/1_head.tex
move from MS word for clean up

Authors
Haoyu Yang, Juanli Zhao, Qiankun Wang, Bin Lin, Wei Luo, Ziqi Sun*, Ting Liao*

Title
Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO2 Reduction Reactions