### 2024-01-04 paper/si_sections/3_others.tex

Section 3 Other additional information
Section 3.1 Additional information on machine learning environment.
The hyperparameter optimization for the proposed CNN model was performed using NVIDIA® V100 Tensor Core GPUs, courtesy of the Australian National Computational Infrastructure's (NCI) HPC Gadi system. The HyperBand algorithm [7], as implemented in KerasTuner [8] , facilitated the tuning process.

Supplementary Table 19
Core packages in deep learning environment
Name	Version	Build	Channel
conda	4.13.0	py39h06a4308_0	anaconda
cuda-nvcc	11.8.89	0	nvidia
cuda-toolkit	11.8.0	0	nvidia
cudatoolkit	11.6.0	habf752d_9	nvidia
ipykernel	6.9.1	py39h06a4308_0	anaconda
ipython	7.34.0	pypi_0	pypi
jupyter_client	7.2.2	py39h06a4308_0	anaconda
jupyterlab	3.4.4	py39h06a4308_0	anaconda
keras	2.9.0	pypi_0	pypi
keras-preprocessing	1.1.2	pypi_0	pypi
keras-tuner	1.3.5		pypi
matplotlib	3.5.1	py39h06a4308_1	anaconda
numpy	1.22.4	pypi_0	pypi
pandas	1.3.4	py39h8c16a72_0	anaconda
python	3.9.12	h12debd9_1	anaconda
scikit-learn	1.1.1	py39h6a678d5_0	anaconda
scipy	1.8.1	py39hddc5342_3	conda-forge
tensorboard	2.9.1	pypi_0	pypi
tensorflow	2.9.3	pypi_0	pypi
tensorflow-gpu	2.9.3	pypi_0	pypi
xlrd	2.0.1	pyhd3eb1b0_0	anaconda
yaml	0.2.5	h7b6447c_0	anaconda
Note. The complete list of packages is provided along with the source code.

Section 3.2 Additional information on data analysis and visualization environment.
Data analysis and visualization tasks were executed on an Apple MacBook Air featuring an octa-core M2 chip. For accuracy, the TensorFlow-Metal plugin for Mac-based GPU acceleration was deliberately excluded, as it could potentially yield erroneous CNN model predictions.

The illustrations of CNN architecture displayed in Figure 1 of the main text, as well as  Supplementary Figure 21 and Supplementary Figure 30, were obtained from the GitHub repository dair-ai/ml-visuals [9]. Blender was used to create and render the visual representation of the catalyst analysis pipeline shown in Figure 1 of the main text.

Supplementary Table 20
Core packages in data analysis and visualization environment
Name	Version	Build	Channel
bokeh	3.2.2	pypi_0	pypi
colorcet	3.0.1	pypi_0	pypi
heatmapz	0.0.4	pypi_0	pypi
keras	2.13.1	pypi_0	pypi
keras-tuner	1.3.5	pypi_0	pypi
matplotlib	3.7.2	pypi_0	pypi
numpy	1.24.3	pypi_0	pypi
pandas	2.0.3	pypi_0	pypi
python	3.11.4	hb885b13_0
pyyaml	6.0.1	pypi_0	pypi
scikit-learn	1.3.0	pypi_0	pypi
scipy	1.11.1	pypi_0	pypi
seaborn	0.12.2	pypi_0	pypi
tensorboard	2.13.0	pypi_0	pypi
tensorflow	2.13.0	pypi_0	pypi
tensorflow-macos	2.13.0	pypi_0	pypi
Note. The complete list of packages is provided along with the source code.