### 2024-01-04 paper/main_sections/1_head.tex
temp save: format cleanup

% Title and Authors (Affiliations)

% Title
\title{Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO$_2$ Reduction Reactions}

% Authors and Affiliations
\author[1]{Haoyu Yang}
\author[3]{Juanli Zhao}
\author[4]{Qiankun Wang}
\author[5]{Bin Lin}
\author[6]{Wei Luo}
\author[1,*]{Ziqi Sun}
\author[2,**]{Ting Liao}

\affil[1]{School of Chemistry and Physics, Queensland University of Technology, George Street, Brisbane, QLD 4000, Australia}
\affil[2]{School of Mechanical Medical and Process Engineering, Queensland University of Technology, George Street, Brisbane, QLD 4000, Australia}
\affil[3]{School of Materials Science and Engineering, Shanghai University, Shanghai, China}
\affil[4]{School of Computer Science and Engineering, Southeast University, Nanjing, China}
\affil[5]{State Key Laboratory for Modification of Chemical Fibers and Polymer Materials, College of Materials Science and Engineering, Institute of Functional Materials, Donghua University, Shanghai, 201620 China}
\affil[*]{Corresponding author: ziqi.sun@qut.edu.au}
\affil[**]{Corresponding author: t3.liao@qut.edu.au}

% Date (optional)
\date{2024-01-04}

% Make Title
\maketitle