### 2024-01-05 paper/si_sections/3_others.tex

% Supporting Information Section Three: Other Additional Information

\section{Other additional information}

\subsection{Additional information on machine learning environment}

The hyperparameter optimization for the proposed CNN model was performed using NVIDIA$^{\textregistered}$ (TODO:) V100 Tensor Core GPUs, courtesy of the Australian National Computational Infrastructure's (NCI) HPC Gadi system. The HyperBand algorithm \cite{li2018hyperband}, as implemented in KerasTuner \cite{omalley2019kerastuner} , facilitated the tuning process.

% Supplementary Table 19
\begin{table}[h]
    \centering
    \begin{tabular}{lr}
      \hline
      Name	Version	Build	Channel
      \hline
      conda	4.13.0	py39h06a4308_0	anaconda
      cuda-nvcc	11.8.89	0	nvidia
      cuda-toolkit	11.8.0	0	nvidia
      cudatoolkit	11.6.0	habf752d_9	nvidia
      ipykernel	6.9.1	py39h06a4308_0	anaconda
      ipython	7.34.0	pypi_0	pypi
      jupyter_client	7.2.2	py39h06a4308_0	anaconda
      jupyterlab	3.4.4	py39h06a4308_0	anaconda
      keras	2.9.0	pypi_0	pypi
      keras-preprocessing	1.1.2	pypi_0	pypi
      keras-tuner	1.3.5		pypi
      matplotlib	3.5.1	py39h06a4308_1	anaconda
      numpy	1.22.4	pypi_0	pypi
      pandas	1.3.4	py39h8c16a72_0	anaconda
      python	3.9.12	h12debd9_1	anaconda
      scikit-learn	1.1.1	py39h6a678d5_0	anaconda
      scipy	1.8.1	py39hddc5342_3	conda-forge
      tensorboard	2.9.1	pypi_0	pypi
      tensorflow	2.9.3	pypi_0	pypi
      tensorflow-gpu	2.9.3	pypi_0	pypi
      xlrd	2.0.1	pyhd3eb1b0_0	anaconda
      yaml	0.2.5	h7b6447c_0	anaconda
      \hline
    \end{tabular}
    \caption{Core packages in deep learning environment}
    \label{si_table19}
\end{table}
Note. The complete list of packages is provided along with the source code.

\subsection{Additional information on data analysis and visualization environment}
Data analysis and visualization tasks were executed on an Apple MacBook Air featuring an octa-core M2 chip. For accuracy, the TensorFlow-Metal plugin for Mac-based GPU acceleration was deliberately excluded, as it could potentially yield erroneous CNN model predictions.

The illustrations of CNN architecture displayed in Figure 1 of the main text, as well as  Supplementary Figure 21 and Supplementary Figure 30, were obtained from the GitHub repository dair-ai/ml-visuals \cite{Saravia_ML_Visuals_2021}. Blender was used to create and render the visual representation of the catalyst analysis pipeline shown in Figure 1 of the main text.

% Supplementary Table 20
\begin{table}[h]
    \centering
    \begin{tabular}{lr}
      \hline
      Name	Version	Build	Channel
      \hline
      bokeh	3.2.2	pypi_0	pypi
      colorcet	3.0.1	pypi_0	pypi
      heatmapz	0.0.4	pypi_0	pypi
      keras	2.13.1	pypi_0	pypi
      keras-tuner	1.3.5	pypi_0	pypi
      matplotlib	3.7.2	pypi_0	pypi
      numpy	1.24.3	pypi_0	pypi
      pandas	2.0.3	pypi_0	pypi
      python	3.11.4	hb885b13_0
      pyyaml	6.0.1	pypi_0	pypi
      scikit-learn	1.3.0	pypi_0	pypi
      scipy	1.11.1	pypi_0	pypi
      seaborn	0.12.2	pypi_0	pypi
      tensorboard	2.13.0	pypi_0	pypi
      tensorflow	2.13.0	pypi_0	pypi
      tensorflow-macos	2.13.0	pypi_0	pypi
      \hline
    \end{tabular}
    \caption{Core packages in data analysis and visualization environment}
    \label{si_table20}
\end{table}
Note. The complete list of packages is provided along with the source code.