### 2024-01-19 paper/arxiv/main_1_head.tex
fix affi order

% Title and Authors (Affiliations)

% Title
\title{Convolutional Neural Networks and Volcano Plots:
Screening and Prediction of Two-Dimensional Single-Atom Catalysts
for CO\textsubscript{2} Reduction Reactions}

% Authors
\author[1]{Haoyu Yang}
\author[2]{Juanli Zhao}
\author[3]{Qiankun Wang}
\author[2]{Bin Liu}
\author[4]{Wei Luo}
\author[1,*]{Ziqi Sun}
\author[5,*]{Ting Liao}

% Affiliations
\affil[1]{School of Chemistry and Physics, Queensland University of Technology,
George Street, Brisbane, QLD 4000, Australia}
\affil[2]{School of Materials Science and Engineering, Shanghai University, Shanghai, China}
\affil[3]{School of Computer Science and Engineering, Southeast University, Nanjing, China}
\affil[4]{College of Materials Science and Engineering, Institute of Functional Materials,
Donghua University, Shanghai, China}
\affil[5]{School of Mechanical Medical and Process Engineering, Queensland University of Technology,
George Street, Brisbane, QLD 4000, Australia}
\affil[*]{Corresponding author: ziqi.sun@qut.edu.au}
\affil[*]{Corresponding author: t3.liao@qut.edu.au}

% Date
\date{January 19, 2024}

\maketitle