### 2024-01-27 paper/arxiv-submission/main.tex
copy for cleanup: arxiv submit

% Journal Paper Main Text Header

% Document Class and Packages
\documentclass[a4paper, 12pt, titlepage]{article}

% Import packages
\usepackage{amsmath}
\usepackage{array}  % Required for m{} column type
\usepackage{bookmark}
\usepackage{booktabs}  % For better horizontal lines
\usepackage{caption}  % To customize captions
\usepackage{cleveref}  % Define Figure/Table cross-ref headers
\usepackage{datetime2}
\usepackage{enumitem}
\usepackage{geometry}  % Adjust page geometry if needed
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{nicematrix}
\usepackage{tablefootnote}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xr}  % external cross-ref
\usepackage[affil-it]{authblk}
\usepackage[english]{babel}
\usepackage[version=4]{mhchem}

% Define Figure/Table/Equation cross-reference headers
\crefname{figure}{Figure}{Figures}
\crefname{table}{Table}{Tables}
\crefname{equation}{Equation}{Equations}
\captionsetup{labelfont=bf}  % make caption label bold

\externaldocument{supp}  % for cross-ref

\begin{document}
\pdfoutput=1  % declare PDFLaTeX processing

\newpage
\input{main_1_head.tex}  % Title and Author Information
\newpage
\input{main_2_abstract.tex}
\newpage
\input{main_3_introduction.tex}
\newpage
\input{main_4.1_cnn_for_eads.tex}
\input{main_4.2_volcano_plots.tex}
\input{main_4.3_validation.tex}
\input{main_4.4_shifting_exp.tex}
\newpage
\input{main_5_methods.tex}

\appendix
\newpage
\input{main_6_appendix.tex}  % Appendix

% Bibliography
\newpage
\bibliographystyle{plain}
\bibliography{references}

\end{document}

% Notes:
% Table formatted to APA format following: https://tex.stackexchange.com/questions/681466/annotation-for-table-following-the-apa-style