### 2024-01-27 paper/arxiv-submission/main.tex
remove comments from main

\documentclass[a4paper, 12pt, titlepage]{article}

\usepackage{amsmath}
\usepackage{array}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{cleveref}
\usepackage{datetime2}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{nicematrix}
\usepackage{tablefootnote}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xr}
\usepackage[affil-it]{authblk}
\usepackage[english]{babel}
\usepackage[version=4]{mhchem}

\crefname{figure}{Figure}{Figures}
\crefname{table}{Table}{Tables}
\crefname{equation}{Equation}{Equations}
\captionsetup{labelfont=bf}

\externaldocument{supp}

\begin{document}
\pdfoutput=1

\newpage
\input{main_1_head.tex}
\newpage
\input{main_2_abstract.tex}
\newpage
\input{main_3_introduction.tex}
\newpage
\input{main_4.1_cnn_for_eads.tex}
\input{main_4.2_volcano_plots.tex}
\input{main_4.3_validation.tex}
\input{main_4.4_shifting_exp.tex}
\newpage
\input{main_5_methods.tex}

\appendix
\newpage
\input{main_6_appendix.tex}

\newpage
\bibliographystyle{plain}
\bibliography{references}

\end{document}

\typeout{get arXiv to do 4 passes: Label(s) may have changed. Rerun}