### 2024-02-09 paper/journal-submission/cover-letter/cover_letter.tex
add cover letter skeleton

% Cover Letter
\documentclass[a4paper, 12pt]{article}

\usepackage{booktabs}
\usepackage{graphics}
\usepackage{multirow}
\usepackage{tablefootnote}

\begin{document}

\title{Cover Letter for: Convolutional Neural Networks and Volcano Plots:
    Screening and Prediction of Two-Dimensional Single-Atom Catalysts
    for CO\textsubscript{2} Reduction Reactions}

\date{\today}
\maketitle
\newpage

\section{Affiliation and contact information for the corresponding authors}

\begin{table}
    \label{cov_let_table:corr_author_contact}
    \centering
    \begin{tabular}{p{3cm}p{6cm}l}
    \toprule
    Corresponding Author  & Affiliation                                             & Contact Email       \\
    \midrule
    Ting Liao             & School of Mechanical Medical and Process Engineering,
                            Queensland University of Technology, George Street,
                            Brisbane, QLD 4000, Australia                           & t3.liao@qut.edu.au   \\
    Ziqi Sun              & School of Chemistry and Physics,
                            Queensland University of Technology, George Street,
                            Brisbane, QLD 4000, Australia                           & ziqi.sun@qut.edu.au  \\
    \bottomrule
    \end{tabular}
\end{table}

\section{Importance of this work}

\section{Reviewers and exclusion of referees}

TODO: need supervisor input.

\section{Prior discussions with editor}

Authors had no prior discussions with \textit{Nature Communications} editors about this work.

\end{document}
% End of Cover Letter