### 2024-03-20 paper/journal-submission/supplement.tex
switch to pdf for figures

% Supplementary Information

% Title
\begin{center}
  \Large{Supporting Information for:  \\
  Convolutional Neural Networks and Volcano Plots:
  Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO\textsubscript{2} Reduction Reactions}
\end{center}

% SI Section One: Additional details on DFT calculations
\newpage
\section{Additional details on DFT calculations}

\subsection{Catalyst models}
\label{supp_sec2.1_catalysts}

Representative structures for each substrate are depicted in \cref{supp_fig2:Ge_g-C3N4} through \cref{supp_fig7:Al-C2N}. For the adsorption process, the initial state is defined by positioning the optimized adsorbates at a distance of 6.5 \text{\AA} from the single-metal atoms supported on the substrate, oriented along the z-axis.

% SI Figure 1: Periodic table of investigated metals
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig1_ptable.pdf}
  \caption{\textbf{Periodic table of investigated metals.}
  Periodic table highlighting investigated elements: transition metals in blue, metalloids in purple, and other metals in orange.}
  \label{supp_fig1:ptable}
\end{figure}

% SI Figure 2: Structure of Ge atom supported on g-C3N4
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig2_Ge_g-C3N4.pdf}
  \caption{\textbf{Structure of single germanium atom supported on graphitic nitride.}
  (\textbf{a}) View along the z-axis and (\textbf{b}) view along the x-axis
  of the Ge@g-C\textsubscript{3}N\textsubscript{4} structure.
  In the illustrations, silver spheres represent N atoms, brown spheres
  signify C atoms, and purple sphere indicates Ge atom.}
  \label{supp_fig2:Ge_g-C3N4}
\end{figure}

% SI Figure 3: Structure of Cr atom supported on nitrogen-doped graphene
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig3_Cr-n-gra.pdf}
  \caption{\textbf{Structure of single chromium atom supported on nitrogen-doped graphene.}
  (\textbf{a}) View along the z-axis and (\textbf{b}) view along the x-axis
  of the Cr@nitrogen-doped-graphene structure.
  In the illustrations, silver spheres represent N atoms, brown spheres
  signify C atoms, and indigo sphere indicates Cr atom.}
  \label{supp_fig3:Cr-n-gra}
\end{figure}

% SI Figure 4: Structure of Os atom supported on graphene with dual-vacancy
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig4_Os-gra-vac.pdf}
  \caption{\textbf{Structure of single osmium atom supported on graphene with dual-vacancy.}
  (\textbf{a}) View along the z-axis and (\textbf{b}) view along the x-axis
  of the Os@graphene-with-dual-vacancy structure.
  In the illustrations, brown spheres signify C atoms, and light yellow sphere
  indicates Os atom.}
  \label{supp_fig4:Os-gra-vac}
\end{figure}

% SI Figure 5: Structure of Cr atom supported on black phosphorous
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig5_Cr-BP.pdf}
  \caption{\textbf{Structure of single chromium atom supported on black phosphorous.}
  (\textbf{a}) View along the z-axis and (\textbf{b}) view along the x-axis
  of the Cr@black phosphorous structure.
  In the illustrations, light purple spheres represent P atoms, and indigo sphere
  indicates Cr atom.}
  \label{supp_fig5:Cr-BP}
\end{figure}

% SI Figure 6: Structure of Y atom supported on boron nitride
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig6_Y-BN.pdf}
  \caption{\textbf{Structure of single yttrium atom supported on boron nitride.}
  (\textbf{a}) View along the z-axis and (\textbf{b}) view along the x-axis
  of the Y@boron-nitride structure.
  In the illustrations, silver spheres represent N atoms, green spheres
  signify B atoms, and light green sphere indicates Y atom.}
  \label{supp_fig6:Y-BN}
\end{figure}

% SI Figure 7: Structure of Al atom supported on C2N
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig7_Al-C2N.pdf}
  \caption{\textbf{Structure of single aluminum atom supported on C\textsubscript{2}N.}
  (\textbf{a}) View along the z-axis and (\textbf{b}) view along the x-axis
  of the Al@C\textsubscript{2}N structure.
  In the illustrations, silver spheres represent N atoms, brown spheres
  signify C atoms, and blue sphere indicates Al atom.}
  \label{supp_fig7:Al-C2N}
\end{figure}

\subsection[CO2RR pathways]{CO\textsubscript{2}RR pathways}
\label{supp_sec2.2_co2rr_paths}

In this study, we explored three reaction pathways that have been previously documented in the literature \cite{durand2011structure, nie2014reaction, peterson2010copper}, as depicted in \cref{supp_fig8:co2rr_paths}.
The adsorption configuration of the *CHO intermediate is pivotal in each of these pathways for determining the reaction mechanism in the CO\textsubscript{2}RR process.
We calculated the energies associated with the *CHO intermediate for catalysts
supported on g-C\textsubscript{3}N\textsubscript{4}, nitrogen-doped graphene, and dual-vacancy graphene-under each mechanism.
These calculations are summarized in \cref{supp_table1:e_3_co2rr_paths}.
Our results indicate that Pathway 1 is the most energetically favorable for most catalysts we investigated.
Therefore, to streamline our analysis, we focused exclusively on Mechanism 1 across all catalysts examined.

% SI Figure 8: Three investigated CO2RR pathways
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\textwidth]{supp_fig8_co2rr_paths.pdf}
  \caption{\textbf{Three investigated CO\textsubscript{2}RR pathways \cite{durand2011structure, nie2014reaction, peterson2010copper} investigated in this work.}}
  \label{supp_fig8:co2rr_paths}
\end{figure}

% Supplementary Table 1: DFT-calculated energies for protonated *CO intermediate for 3 pathways
\begin{table}[htbp]
\label{supp_table1:e_3_co2rr_paths}
  \caption{DFT-calculated energies in eV for protonated *CO intermediate
    for three investigated CO\textsubscript{2}RR pathways.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{l *{9}{l}}
      \toprule
      \multirow{2}{*}{Metal}
      & \multicolumn{3}{c}{g-C\textsubscript{3}N\textsubscript{4}}
      & \multicolumn{3}{c}{nitrogen-doped graphene}
      & \multicolumn{3}{c}{graphene with dual-vacancy}  \\ \cline{2-10}
      & Pathway-1 & Pathway-2 & Pathway-3
      & Pathway-1 & Pathway-2 & Pathway-3
      & Pathway-1 & Pathway-2 & Pathway-3  \\
      \midrule
      Al & \textbf{-499.3494} & -497.8343 & -499.3005 & \textbf{-670.3489} & -668.5142 & -670.3093 & -667.2538 & -666.6725 & \textbf{-667.2665} \\
      Co & -502.5946 & -501.1257 & \textbf{-502.6139} & -672.8847 & -670.8013 & \textbf{-672.9020} & \textbf{-671.6284} & -670.4388 & -671.3827 \\
      Cr & -505.6012 & -504.2022 & \textbf{-505.6029} & -675.3126 & -674.1640 & \textbf{-675.3480} & \textbf{-673.4283} & -672.1919 & -673.3818 \\
      Cu & \textbf{-495.8913} & -493.6853 & -495.8820 & \textbf{-666.9894} & -664.9810 & -666.9705 & -666.9876 & -664.7684 & \textbf{-667.1395} \\
      Fe & \textbf{-501.3325} & -499.3969 & -501.3221 & -673.8242 & -672.0887 & \textbf{-673.8478} & \textbf{-672.7810} & -671.5694 & -672.4960 \\
      Ga & -498.2365 & -496.3967 & \textbf{-498.2912} & \textbf{-667.3963} & -665.4677 & -667.3482 & -665.7780 & -664.8573 & \textbf{-665.8030} \\
      Ge & -500.2052 & -497.9780 & \textbf{-500.2156} & \textbf{-668.1785} & -666.3484 & -668.1463 & \textbf{-667.7387} & -665.7394 & -667.7324 \\
      Mn & \textbf{-502.6355} & -500.8006 & -502.6294 & -674.6848 & -672.9574 & \textbf{-674.7030} & -673.2243 & -672.6495 & \textbf{-673.2361} \\
      Ni & \textbf{-500.4085} & -498.5165 & -500.3985 & \textbf{-670.0979} & -667.9544 & -670.0832 & -670.1619 & -668.3684 & \textbf{-670.1756} \\
      Sc & -503.4428 & -501.6129 & \textbf{-503.4437} & \textbf{-673.9590} & -671.7624 & -673.3818 & \textbf{-670.0841} & -668.1044 & -669.6050 \\
      Ti & -503.7848 & -502.6148 & \textbf{-503.8033} & \textbf{-675.1006} & -673.2417 & -674.6257 & \textbf{-672.7127} & -670.6787 & -672.2491 \\
      V  & \textbf{-503.6115} & -502.6188 & -503.5956 & \textbf{-675.0533} & -674.0394 & -675.0136 & \textbf{-673.1855} & -671.5382 & -673.1583 \\
      Zn & \textbf{-493.9854} & -491.5010 & -493.9743 & \textbf{-664.9935} & -663.0932 & -664.9758 & -663.7735 & -661.8097 & \textbf{-663.7959} \\
      Ag & \textbf{-494.4055} & -492.4511 & -494.3905 & -664.6322 & -662.9060 & \textbf{-664.6353} & -664.3972 & -662.3588 & \textbf{-665.0570} \\
      Cd & \textbf{-493.3084} & -491.1781 & -493.2951 & \textbf{-663.8832} & -661.6272 & -663.8653 & -661.8518 & -659.9601 & \textbf{-661.8631} \\
      In & -498.5272 & -496.6593 & \textbf{-498.5908} & \textbf{-666.3968} & -663.7707 & -666.3845 & -663.6532 & -661.5581 & \textbf{-663.6913} \\
      Mo & \textbf{-504.5248} & -503.3763 & -504.5148 & -675.3472 & -674.7118 & \textbf{-675.4389} & -674.7186 & -672.9335 & \textbf{-674.7397} \\
      Nb & -505.3967 & -504.7113 & \textbf{-505.3982} & \textbf{-676.0492} & -674.1997 & -675.4107 & \textbf{-674.6606} & -672.8268 & -674.5488 \\
      Pd & -500.0370 & -498.3310 & \textbf{-500.0455} & \textbf{-669.0243} & -667.0759 & -669.0205 & \textbf{-669.1169} & -667.3223 & -669.1119 \\
      Rh & \textbf{-499.2225} & -497.3162 & -499.2170 & -672.7920 & -670.7853 & \textbf{-672.8099} & \textbf{-671.4286} & -670.3036 & -671.4181 \\
      Ru & \textbf{-503.9986} & -502.1908 & -503.9640 & -674.0273 & -672.4060 & \textbf{-674.0339} & \textbf{-672.9956} & -672.1332 & -672.9909 \\
      Sb & \textbf{-497.3836} & -495.1348 & -497.3442 & \textbf{-666.7198} & -664.9611 & -666.7190 & -666.8632 & -664.8131 & \textbf{-666.8670} \\
      Sn & \textbf{-499.4952} & -497.2007 & -499.1116 & \textbf{-667.1459} & -665.3835 & -667.0999 & \textbf{-665.9821} & -663.8277 & -665.9755 \\
      Tc & \textbf{-503.0602} & -501.1754 & -503.0532 & -674.4342 & -673.8164 & \textbf{-674.4559} & -674.0615 & -672.9747 & \textbf{-674.0738} \\
      Y  & \textbf{-503.9232} & -502.0670 & -503.8988 & \textbf{-673.7620} & -671.6132 & -673.2241 & -669.4633 & -668.0403 & \textbf{-669.5030} \\
      Zr & \textbf{-505.2907} & -503.9148 & -505.1817 & \textbf{-675.6874} & -673.5456 & -675.0519 & \textbf{-673.1694} & -671.1993 & -672.8072 \\
      Au & \textbf{-497.0784} & -495.1720 & -497.0175 & \textbf{-665.1479} & -663.3239 & -665.1370 & -666.4185 & -664.2892 & \textbf{-666.5182} \\
      Bi & \textbf{-496.9946} & -494.7679 & -496.9934 & -665.9563 & -664.2409 & \textbf{-666.0512} & \textbf{-665.0896} & -663.2974 & -665.0887 \\
      Hf & \textbf{-506.4837} & -504.9734 & -506.4777 & \textbf{-677.3077} & -675.1797 & -676.6502 & \textbf{-674.7209} & -672.6650 & -674.2149 \\
      Hg & \textbf{-491.0035} & -488.7951 & -490.9728 & \textbf{-661.2458} & -659.2406 & -661.2398 & -661.7192 & -659.8398 & \textbf{-661.7300} \\
      Ir & \textbf{-503.0227} & -501.1368 & -503.0217 & -673.9926 & -672.0068 & \textbf{-674.0087} & \textbf{-673.5129} & -672.6725 & -673.4975 \\
      Os & -504.7648 & -502.9878 & \textbf{-504.8105} & -675.2098 & -673.6653 & \textbf{-675.2235} & \textbf{-674.9483} & -674.1970 & -674.9223 \\
      Pb & \textbf{-499.1183} & -497.1173 & -499.0544 & -666.4446 & -664.6897 & \textbf{-666.4625} & \textbf{-664.0200} & -662.2222 & -664.0123 \\
      Pt & \textbf{-499.3356} & -497.5110 & -499.3215 & -670.3996 & -668.2412 & \textbf{-670.4147} & \textbf{-671.1408} & -669.2725 & -671.1370 \\
      Re & \textbf{-504.1830} & -502.4735 & -504.1746 & \textbf{-675.7670} & -675.0715 & -675.7486 & \textbf{-675.8701} & -674.9245 & -675.8233 \\
      Ta & -506.6180 & -505.7793 & \textbf{-506.6241} & \textbf{-677.7222} & -675.9907 & -677.0911 & \textbf{-676.4371} & -674.5434 & -676.4262 \\
      Tl & \textbf{-497.9559} & -496.1578 & -497.9382 & -664.5047 & -662.7051 & \textbf{-664.5051} & -662.6279 & -660.6086 & \textbf{-662.6317}\\
      W  & \textbf{-505.8134} & -504.9085 & -505.8030 & -676.5304 & -676.1384 & \textbf{-676.9965} & \textbf{-676.7669} & -675.4895 & -676.7356 \\
      \bottomrule
    \end{tabular}
  }

  \smallskip

  \footnotesize\textit{Note:} Energies of the most stable configurations are highlighted in bold.
\end{table}

\subsection{DFT energies and energy corrections}
\label{supp_sec2.3_energies}

% Supplementary Table 2: Free energies for isolated species.
\begin{table}[htbp]
\label{supp_table2:species_free_energies}
  \caption{Free energies at 298.15 K for isolated species.}
  \center
  \small
  \begin{tabularx}{0.5\textwidth}{@{}lXr@{}}
    \toprule
    Species               & Free Energy (eV)  \\
    \midrule
    CO\textsubscript{2}   & -23.3140          \\
    CO                    & -15.3336          \\
    CH\textsubscript{4}   & -23.4639          \\
    H\textsubscript{2}    & -6.9315           \\
    H\textsubscript{2}O   & -14.3239          \\
    COOH                  & -24.3963          \\
    CHO                   & -17.0983          \\
    CH\textsubscript{2}O  & -22.1607          \\
    OCH\textsubscript{3}  & -24.5075          \\
    O                     & -1.9007           \\
    OH                    & -7.7303           \\
    \bottomrule
  \end{tabularx}

  \smallskip

  \begin{flushright}
  \begin{minipage}{\textwidth}
    \footnotesize\textit{Note:} VASPKIT \cite{wang2021vaspkit} was employed to calculate free energies.
  \end{minipage}
  \end{flushright}
\end{table}

% Supplementary Table 3: Free energy corrections for adsorbed intermediates
\begin{table}[htbp]
\label{supp_table3:ads_e_correction}
  \caption{Free energy corrections at 298.15 K for adsorbed intermediates.}
  \centering
  \small
  \begin{tabularx}{0.5\textwidth}{@{}lXr@{}}
    \toprule
    Species                & Free energy correction (eV)  \\
    \midrule
    *CO\textsubscript{2}   &  0.0864                      \\
    *COOH                  &  0.3693                      \\
    *CO                    &  0.0111                      \\
    *CHO                   &  0.2716                      \\
    *CH\textsubscript{2}O  &  0.5595                      \\
    *OCH\textsubscript{3}  &  0.8738                      \\
    *O                     & -0.0193                      \\
    *OH                    &  0.1993                      \\
    *H                     &  0.1533                      \\
    \bottomrule
  \end{tabularx}

  \smallskip

  \begin{flushright}  % justify footnote
  \begin{minipage}{\textwidth}
    \footnotesize\textit{Note:} {Free energy corrections were obtained by averaging the
      DFT-computed corrections, which include zero-point energy and entropy, across three substrates:
      g-C\textsubscript{3}N\textsubscript{4}, nitrogen-doped graphene, and dual-vacancy graphene.}
  \end{minipage}
  \end{flushright}
\end{table}

% Supplementary Table 4: DFT calculated final state energies for adsorbates supported on g-c3n4
\begin{table}[htbp]
\label{supp_table4:e_g-c3n4}
  \caption{DFT calculated final state energies in eV for adsorbates
    supported on g-C\textsubscript{3}N\textsubscript{4}.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{l *{9}{l}}
      \toprule
      Metal & *CO\textsubscript{2} & *COOH    & *CO       & *CHO
      & *CH\textsubscript{2}O  & *OCH\textsubscript{3}  & *O        & *OH       & *H        \\
      \midrule
      Al & -481.2292 & -504.7972 & -508.6897 & -496.4490 & -500.7640 & -503.7935 & -509.9998 & -488.6623 & -485.4855 \\
      Co & -483.2039 & -507.0765 & -510.0453 & -500.1329 & -503.9316 & -507.0552 & -510.9052 & -489.9487 & -486.6194 \\
      Cr & -486.7813 & -509.9599 & -513.3962 & -502.1493 & -505.7125 & -509.8851 & -514.1702 & -493.7457 & -489.9202 \\
      Cu & -477.5955 & -502.5914 & -505.9613 & -495.6511 & -498.0502 & -502.4590 & -506.2829 & -484.3166 & -481.9683 \\
      Fe & -482.0576 & -507.1508 & -510.2918 & -500.1101 & -502.1626 & -506.0115 & -511.2543 & -490.0116 & -486.4067 \\
      Ga & -480.8854 & -503.9317 & -506.4938 & -495.7379 & -498.7813 & -503.1345 & -507.5029 & -486.1702 & -483.4575 \\
      Ge & -481.5807 & -504.6467 & -508.4204 & -496.4421 & -500.5712 & -504.0218 & -509.4116 & -488.4168 & -485.2680 \\
      Mn & -483.7690 & -508.5410 & -511.6514 & -500.6279 & -502.6739 & -508.6433 & -512.8569 & -491.2239 & -488.1082 \\
      Ni & -482.0061 & -505.1988 & -508.4173 & -498.5054 & -501.0040 & -504.8945 & -508.6266 & -487.3684 & -484.8272 \\
      Sc & -483.9963 & -509.3147 & -512.9882 & -501.4505 & -505.2024 & -509.4054 & -514.5969 & -493.4105 & -489.2632 \\
      Ti & -484.2282 & -509.6510 & -512.8528 & -502.0410 & -505.3761 & -509.9729 & -514.0700 & -493.6478 & -489.2234 \\
      V  & -484.1990 & -508.6926 & -512.3009 & -501.6777 & -504.4297 & -509.3618 & -513.4561 & -493.8569 & -488.9474 \\
      Zn & -475.2375 & -500.3556 & -504.4103 & -492.4144 & -496.5426 & -499.4232 & -504.6297 & -482.6782 & -481.0438 \\
      Ag & -476.7049 & -501.6430 & -504.5861 & -493.9138 & -496.7220 & -500.9477 & -504.5408 & -482.6604 & -481.2659 \\
      Cd & -475.1288 & -500.1854 & -503.4108 & -491.8954 & -495.6486 & -499.4959 & -503.6209 & -483.3427 & -479.8278 \\
      In & -481.1330 & -504.1653 & -506.4422 & -496.0016 & -498.6643 & -503.3286 & -507.3780 & -485.6281 & -483.1445 \\
      Mo & -484.2892 & -509.4263 & -512.9103 & -503.0025 & -505.5534 & -509.6730 & -513.0136 & -493.5700 & -489.4994 \\
      Nb & -485.4055 & -510.7952 & -514.0380 & -503.5663 & -506.3671 & -511.1572 & -514.7845 & -495.5118 & -490.2111 \\
      Pd & -481.5399 & -504.7593 & -508.5173 & -498.0817 & -501.1620 & -504.0892 & -508.1976 & -486.1551 & -483.9553 \\
      Rh & -479.8884 & -505.2422 & -508.6754 & -498.7129 & -501.4594 & -504.5254 & -508.9536 & -487.8771 & -485.0738 \\
      Ru & -483.8526 & -507.5120 & -511.5881 & -500.6857 & -504.1497 & -506.8934 & -510.9524 & -490.6690 & -488.0169 \\
      Sb & -478.5169 & -502.9708 & -506.5211 & -494.8740 & -498.9453 & -502.5347 & -507.2111 & -486.4448 & -483.4418 \\
      Sn & -481.4238 & -504.4677 & -507.8470 & -496.2883 & -499.9258 & -503.6685 & -508.8152 & -487.6963 & -484.6351 \\
      Tc & -482.7091 & -507.9652 & -512.1064 & -501.9896 & -504.6034 & -508.3486 & -511.4777 & -492.0381 & -488.0897 \\
      Y  & -484.6612 & -509.8564 & -513.6851 & -501.9522 & -505.8238 & -509.8344 & -515.2400 & -493.4038 & -489.6506 \\
      Zr & -485.4694 & -511.1852 & -514.3901 & -503.1958 & -506.9346 & -511.5864 & -515.6720 & -496.3037 & -490.4785 \\
      Au & -479.1941 & -502.3245 & -506.9013 & -495.6782 & -499.0711 & -501.6877 & -506.5462 & -484.6198 & -483.6909 \\
      Bi & -478.2695 & -502.8490 & -506.2055 & -494.6965 & -498.5557 & -502.3398 & -506.7999 & -485.7536 & -483.1348 \\
      Hf & -486.5155 & -512.4535 & -515.5965 & -504.3393 & -508.1599 & -512.8332 & -517.2187 & -496.4195 & -491.7486 \\
      Hg & -472.8730 & -500.0491 & -502.1889 & -491.7544 & -494.1224 & -499.2181 & -501.8365 & -480.3761 & -479.2046 \\
      Ir & -483.7878 & -507.2024 & -510.6763 & -500.2051 & -503.2948 & -506.5209 & -509.8509 & -490.9790 & -487.3503 \\
      Os & -484.9279 & -508.6427 & -512.5954 & -501.8541 & -505.2080 & -507.8334 & -511.8734 & -491.0155 & -489.0898 \\
      Pb & -481.3587 & -504.4450 & -507.5634 & -496.2379 & -499.6114 & -503.5822 & -508.1698 & -486.8476 & -484.2646 \\
      Pt & -481.1676 & -504.3670 & -508.2170 & -499.0073 & -500.9446 & -504.2695 & -508.1465 & -487.3508 & -484.6644 \\
      Re & -483.6421 & -508.9474 & -512.8388 & -502.8952 & -505.3651 & -509.1168 & -512.6374 & -493.0621 & -489.2811 \\
      Ta & -486.3299 & -512.1371 & -515.1895 & -504.6585 & -507.5888 & -512.5295 & -516.0165 & -497.0134 & -490.5695 \\
      Tl & -480.7357 & -503.7632 & -505.9937 & -495.5894 & -497.7724 & -503.0564 & -506.7919 & -484.0732 & -482.3846 \\
      W  & -485.2596 & -510.8145 & -514.28   & -504.2714 & -506.7988 & -511.0643 & -514.2358 & -494.4330 & -490.6506 \\
      \bottomrule
    \end{tabular}
  }
\end{table}

% Supplementary Table 5: DFT calculated final state energies for adsorbates supported on n-doped graphene
\begin{table}[htbp]
\label{supp_table5:e_n_gra}
    \caption{DFT calculated final state energies in eV for adsorbates supported on nitrogen-doped graphene.}
    \resizebox{\textwidth}{!}{%
      \begin{tabular}{l *{9}{l}}
        \toprule
        Metal & *CO\textsubscript{2} & *COOH    & *CO       & *CHO
        & *CH\textsubscript{2}O  & *OCH\textsubscript{3}  & *O        & *OH       & *H        \\
        \midrule
        Al	& -652.7589	&-675.7997	&-679.7595	&-668.0296	&-671.7596	&-676.2161	&-681.1311	&-658.7849	&-656.4702 \\
        Co	& -654.3280 &-677.5394	&-681.2518	&-670.0860  &-673.9208	&-676.7761	&-680.7219	&-659.1714	&-657.8777 \\
        Cr	& -657.0479	&-680.2074	&-683.7982	&-673.0760  &-676.1859	&-679.4701	&-684.5295	&-664.7306	&-660.4141 \\
        Cu	& -650.2181	&-673.3928	&-675.5873	&-665.2174	&-668.0662	&-672.5797	&-675.7047	&-653.5770  &-652.1010 \\
        Fe	& -655.4034	&-678.5897	&-682.1186	&-671.9101	&-674.7438	&-677.8515	&-682.1216	&-661.5422	&-658.7383 \\
        Ga	& -649.9563	&-672.9807	&-676.9763	&-664.9727	&-669.0739	&-672.8536	&-677.7169	&-655.5345	&-653.9243 \\
        Ge	& -651.0391	&-674.1655	&-677.2582	&-665.9255	&-669.5282	&-673.3438	&-678.1585	&-657.4964	&-654.3444 \\
        Mn	& -656.5635	&-679.7128	&-683.1092	&-672.6753	&-675.6877	&-679.0710  &-683.3898	&-663.1584	&-659.7760 \\
        Ni	& -653.1461	&-676.3270  &-678.5166	&-668.1448	&-671.1502	&-675.5454	&-678.5411	&-656.3929	&-655.1238 \\
        Sc	& -655.2274	&-678.9985	&-682.8810  &-670.9489	&-674.9850  &-679.1986	&-684.3192	&-662.2442	&-659.0243 \\
        Ti	& -655.5937	&-680.2412	&-683.5514	&-671.9689	&-675.9724	&-680.4589	&-685.1366	&-665.2871	&-659.7132 \\
        V	  & -656.0751	&-680.2584	&-683.5739	&-672.6281	&-676.0229	&-680.5926	&-684.9400	&-665.6038	&-659.9017 \\
        Zn	& -648.2316	&-671.3824	&-674.1973	&-663.2623	&-666.3990 	&-670.5613	&-674.5436	&-652.2900  &-650.9227 \\
        Ag	& -646.9759	&-670.2666	&-673.2766	&-663.1058	&-665.7601	&-669.7801	&-673.0113	&-651.0760  &-649.9287 \\
        Cd	& -646.1687	&-669.2620  &-672.8043	&-661.4439	&-664.9473	&-668.8180  &-672.9588	&-650.7684	&-649.7670 \\
        In	& -648.6726	&-671.7218	&-675.3788	&-663.4885	&-667.5183	&-670.7536	&-675.8630  &-653.8535	&-652.4757 \\
        Mo	& -655.7944	&-680.2966	&-683.6262	&-672.7667	&-676.3955	&-680.6291	&-684.8234	&-665.8138	&-659.8169 \\
        Nb	& -656.0277	&-680.8814	&-684.2662	&-672.5576	&-676.9139	&-681.4203	&-685.7720  &-666.2363	&-660.3734 \\
        Pd	& -652.2091	&-675.3967	&-677.2760  &-667.1855	&-669.8609	&-674.6033	&-677.3498	&-655.1080  &-654.0027 \\
        Rh	& -653.6928	&-676.9529	&-680.9755	&-669.5137	&-673.6394	&-676.2879	&-680.2104	&-658.6953	&-657.6908 \\
        Ru	& -654.5425	&-678.2124	&-681.9561	&-672.0589	&-674.6177	&-678.0945	&-681.9008	&-661.3838	&-658.6572 \\
        Sb	& -649.9262	&-672.9182	&-675.6457	&-664.7128	&-667.9159	&-672.0849	&-675.8420 	&-655.1699	&-652.4240 \\
        Sn	& -650.5072	&-673.6118	&-675.7767	&-665.3842	&-668.0544	&-672.7597	&-676.5046	&-655.8947	&-652.9261 \\
        Tc	& -655.4025	&-679.3044	&-682.7454	&-672.5785	&-676.3294	&-679.4840  &-683.3209	&-664.2323	&-659.2703 \\
        Y	  & -655.2660 &-679.0430	&-682.7571	&-670.8208	&-674.7780  &-679.0775	&-684.1210  &-661.8034	&-658.8885 \\
        Zr	& -655.9118	&-680.6859	&-684.1920  &-672.1497	&-676.4014	&-680.7930  &-685.7940  &-665.4878	&-660.4163 \\
        Au	& -648.3344	&-671.3969	&-673.1469	&-663.1955	&-665.5964	&-670.5800	&-673.6092	&-651.2436	&-649.7275 \\
        Bi	& -649.1484	&-672.1225	&-675.1630  &-663.9594	&-667.2981	&-671.4617	&-675.6926	&-654.7684	&-651.8907 \\
        Hf	& -657.3364	&-682.3019	&-685.8119	&-673.6603	&-678.0090  &-682.4001	&-687.4607	&-666.9550  &-662.0651 \\
        Hg	& -644.7368	&-667.9224	&-671.2200	&-659.6906	&-663.5208	&-667.0466	&-670.9016	&-648.8944	&-648.3755 \\
        Ir	& -654.8564	&-678.1079	&-682.2865	&-670.9826	&-674.9484	&-677.4097	&-681.3088	&-660.1543	&-658.9895 \\
        Os	& -655.6722	&-679.5364	&-683.3240  &-673.4289	&-675.9254	&-679.4423	&-683.1300	&-663.0791	&-659.9901 \\
        Pb	& -649.8843	&-672.9588	&-674.5856	&-664.7553	&-667.1174	&-672.0886	&-675.8823	&-653.6827	&-651.5940 \\
        Pt	& -653.3281	&-676.5221	&-678.6094	&-668.2951	&-671.2463	&-675.7317	&-678.4334	&-656.4419	&-655.4733 \\
        Re	& -656.3492	&-680.7664	&-683.9596	&-673.9732	&-676.5164	&-681.0627	&-684.8096	&-665.9526	&-660.5620 \\
        Ta	& -657.1988	&-682.6048	&-685.8231	&-674.1666	&-678.5029	&-682.9709	&-687.2294	&-667.7321	&-661.8794 \\
        Tl	& -648.1346	&-671.1831	&-673.2953	&-662.9610  &-665.6240  &-670.3512	&-673.4567	&-651.7909	&-650.4750 \\
        W	  & -657.0204	&-682.0186	&-685.3735	&-674.3843	&-678.2483	&-682.4377	&-686.3613	&-667.4547	&-661.5686 \\
        \bottomrule
      \end{tabular}
    }
\end{table}

% Supplementary Table 6: DFT calculated final state energies for adsorbates supported on graphene with dual-vac
\begin{table}[htbp]
\label{supp_table6:e_vac_gra}
  \caption{DFT calculated final state energies in eV for adsorbates supported on graphene with dual-vacancy.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{l *{9}{c}}
      \toprule
      Metal & *CO\textsubscript{2} & *COOH    & *CO       & *CHO
      & *CH\textsubscript{2}O  & *OCH\textsubscript{3}  & *O        & *OH       & *H        \\
      \midrule
      Al & -650.3496 & -673.3468 & -676.2273 & -665.6303 & -668.4112 & -672.5763 & -677.4453 & -655.0864 & -653.0166 \\
      Co & -652.5157 & -675.5864 & -679.3861 & -669.0614 & -672.8955 & -675.8508 & -679.5268 & -658.7929 & -655.9377 \\
      Cr & -655.0421 & -678.1256 & -681.8470 & -671.1282 & -674.2868 & -679.1554 & -683.4111 & -663.3554 & -658.4913 \\
      Cu & -649.8187 & -672.9161 & -674.9327 & -664.8190 & -669.0819 & -672.1515 & -674.7438 & -652.8789 & -653.6953 \\
      Fe & -653.7964 & -677.0127 & -680.3412 & -670.0219 & -673.9525 & -676.7631 & -681.1346 & -660.9239 & -656.9759 \\
      Ga & -649.0340 & -672.0433 & -674.9630 & -663.8505 & -667.2101 & -671.2541 & -675.5229 & -653.2086 & -651.8693 \\
      Ge & -651.0133 & -674.2532 & -677.2272 & -666.0554 & -669.5660 & -673.4485 & -677.8152 & -655.7843 & -654.2731 \\
      Mn & -654.8183 & -678.0584 & -681.5834 & -670.9593 & -674.0994 & -677.9260 & -682.6048 & -662.5702 & -658.1064 \\
      Ni & -651.5295 & -674.6692 & -677.8941 & -667.2937 & -670.7325 & -673.9392 & -677.5204 & -656.1026 & -655.6559 \\
      Sc & -652.1630 & -675.5777 & -678.6101 & -667.5024 & -670.7255 & -675.4361 & -680.0287 & -658.0834 & -654.6913 \\
      Ti & -654.2548 & -677.9326 & -681.1516 & -670.0780 & -673.2830 & -677.9113 & -682.6933 & -661.1957 & -657.4040 \\
      V  & -654.6939 & -678.0200 & -681.8606 & -670.8517 & -674.2298 & -679.2097 & -683.4928 & -663.1639 & -658.4619 \\
      Zn & -646.9805 & -670.1676 & -672.2508 & -662.0849 & -664.7199 & -669.3791 & -672.6028 & -650.3134 & -648.8961 \\
      Ag & -647.4840 & -670.5954 & -672.4765 & -662.3757 & -665.1014 & -669.8199 & -672.2893 & -650.2207 & -651.5223 \\
      Cd & -645.2038 & -668.3897 & -670.2685 & -660.2502 & -662.7479 & -667.5994 & -670.6023 & -648.2958 & -646.9658 \\
      In & -646.8189 & -669.9379 & -673.1820 & -661.7351 & -665.3833 & -669.1524 & -673.5557 & -651.3730 & -650.2004 \\
      Mo & -656.2058 & -679.7202 & -683.2424 & -672.3832 & -675.6459 & -679.9903 & -684.8002 & -664.6331 & -659.9673 \\
      Nb & -656.1010 & -679.4842 & -683.3623 & -672.0587 & -675.6683 & -679.5996 & -684.8935 & -664.2634 & -659.9617 \\
      Pd & -650.5405 & -673.6524 & -677.0119 & -666.0816 & -669.7529 & -672.9394 & -676.5341 & -655.1586 & -654.8764 \\
      Rh & -652.7998 & -675.8992 & -679.6974 & -669.1468 & -672.2062 & -676.2142 & -679.9592 & -659.4214 & -656.7388 \\
      Ru & -654.8166 & -678.0140 & -681.6090 & -671.1415 & -674.0502 & -677.9828 & -682.4415 & -662.4653 & -658.2092 \\
      Sb & -649.9882 & -673.0639 & -675.9906 & -664.8414 & -668.3621 & -672.2685 & -676.6274 & -655.1055 & -653.1345 \\
      Sn & -648.5997 & -671.7246 & -675.2765 & -663.7120 & -667.5399 & -670.8850 & -675.8484 & -653.8138 & -652.4041 \\
      Tc & -655.8173 & -679.2476 & -682.6345 & -672.1709 & -675.0076 & -679.2729 & -683.7569 & -664.1434 & -659.2908 \\
      Y  & -652.1517 & -675.4891 & -678.6682 & -667.4644 & -670.6934 & -675.1478 & -680.0013 & -658.3045 & -654.7163 \\
      Zr & -654.8760 & -678.4572 & -681.6839 & -670.6090 & -673.8291 & -678.4537 & -683.1390 & -661.4741 & -657.9892 \\
      Au & -649.3963 & -672.5116 & -674.3660 & -664.2836 & -667.3357 & -671.7273 & -674.1359 & -652.0358 & -653.1138 \\
      Bi & -648.8651 & -672.0004 & -673.7920 & -663.7934 & -666.2170 & -671.1801 & -674.0672 & -652.8628 & -650.9149 \\
      Hf & -656.1155 & -679.7805 & -683.2130 & -671.9925 & -675.3123 & -679.9256 & -684.7502 & -662.8630 & -659.4860 \\
      Hg & -645.0948 & -668.2801 & -669.9749 & -660.1062 & -662.5696 & -667.4822 & -670.2672 & -647.8582 & -646.7652 \\
      Ir & -654.4999 & -677.8310 & -681.7312 & -671.1936 & -674.7821 & -678.1749 & -681.9532 & -661.9662 & -658.5262 \\
      Os & -656.3635 & -679.8433 & -683.4557 & -673.1036 & -675.9479 & -679.6710 & -684.4704 & -664.7281 & -660.0873 \\
      Pb & -647.8007 & -670.9491 & -673.2996 & -662.7392 & -665.6502 & -670.1373 & -673.5969 & -651.6692 & -650.4432 \\
      Pt & -652.4961 & -675.6092 & -678.9183 & -668.2053 & -671.6867 & -674.8876 & -678.4355 & -657.4936 & -656.2759 \\
      Re & -657.2440 & -681.0777 & -684.5183 & -673.9603 & -676.8950 & -681.2383 & -685.8328 & -666.1615 & -661.2885 \\
      Ta & -657.4068 & -681.0711 & -685.1062 & -673.5947 & -677.3738 & -681.6985 & -686.7561 & -666.0327 & -661.7361 \\
      Tl & -645.7286 & -668.8734 & -671.7205 & -660.7259 & -664.0001 & -668.1110 & -672.0305 & -649.7198 & -648.7563 \\
      W  & -657.5901 & -681.6722 & -685.1010 & -674.1028 & -677.7828 & -682.4493 & -686.6385 & -666.9364 & -661.8641 \\
      \bottomrule
    \end{tabular}
  }
\end{table}

% Supplementary Table 7: ZPE for adsorbates supported on g-c3n4
\begin{table}[htbp]
\label{supp_table7:zpe_g-c3n4}
  \caption{Zero-point energies in eV for relaxed adsorbates supported
    on g-C\textsubscript{3}N\textsubscript{4} at 298.15 K.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{l *{9}{l}}
      \toprule
      Metal & *CO\textsubscript{2} & *COOH & *CO & *CHO
      & *CH\textsubscript{2}O & *OCH\textsubscript{3} & *O & *OH    & *H     \\
      \midrule
      Al & 0.2905 & 0.6130 & 0.1743 & 0.4606 & 0.7456 & 1.0974 & 0.0629 & 0.3521 & 0.1925 \\
      Co & 0.3038 & 0.5895 & 0.2051 & 0.4186 & 0.6269 & 1.0663 & 0.0571 & 0.3248 & 0.1507 \\
      Cr & 0.3140 & 0.5964 & 0.1640 & 0.4374 & 0.7725 & 1.0779 & 0.0670 & 0.3286 & 0.1626 \\
      Cu & 0.3157 & 0.6016 & 0.2045 & 0.4498 & 0.7797 & 1.0751 & 0.0494 & 0.3332 & 0.1577 \\
      Fe & 0.3088 & 0.5859 & 0.1931 & 0.4437 & 0.7720 & 1.0726 & 0.0599 & 0.3326 & 0.1550 \\
      Ga & 0.3102 & 0.6131 & 0.1448 & 0.4665 & 0.7186 & 1.0680 & 0.0608 & 0.3202 & 0.1813 \\
      Ge & 0.3091 & 0.6069 & 0.1396 & 0.4540 & 0.7269 & 1.0881 & 0.0581 & 0.3506 & 0.1904 \\
      Mn & 0.3022 & 0.5876 & 0.1733 & 0.4333 & 0.7637 & 1.0622 & 0.0687 & 0.3219 & 0.1432 \\
      Ni & 0.3023 & 0.5934 & 0.2074 & 0.4546 & 0.7662 & 1.0623 & 0.0542 & 0.3313 & 0.1611 \\
      Sc & 0.2880 & 0.6039 & 0.1771 & 0.4544 & 0.7473 & 1.0806 & 0.0657 & 0.3156 & 0.1480 \\
      Ti & 0.3044 & 0.6088 & 0.1889 & 0.4664 & 0.7863 & 1.0862 & 0.0796 & 0.3417 & 0.1590 \\
      V  & 0.3081 & 0.6005 & 0.2031 & 0.4408 & 0.7868 & 1.0882 & 0.0792 & 0.3388 & 0.1660 \\
      Zn & 0.3243 & 0.6080 & 0.1729 & 0.4601 & 0.7277 & 1.0832 & 0.0514 & 0.3468 & 0.1726 \\
      Ag & 0.3167 & 0.5993 & 0.1815 & 0.4457 & 0.7264 & 1.0590 & 0.0390 & 0.3304 & 0.1576 \\
      Cd & 0.3242 & 0.6007 & 0.1533 & 0.4502 & 0.7292 & 1.0604 & 0.0983 & 0.3322 & 0.1439 \\
      In & 0.3116 & 0.5687 & 0.1399 & 0.4025 & 0.7158 & 1.0551 & 0.0523 & 0.3315 & 0.1141 \\
      Mo & 0.3109 & 0.6071 & 0.2163 & 0.4441 & 0.7970 & 1.0825 & 0.0765 & 0.3312 & 0.1851 \\
      Nb & 0.3101 & 0.5971 & 0.2013 & 0.4340 & 0.7960 & 1.0908 & 0.0798 & 0.3329 & 0.1598 \\
      Pd & 0.3148 & 0.6091 & 0.1996 & 0.4778 & 0.7402 & 1.0777 & 0.0525 & 0.3425 & 0.1508 \\
      Rh & 0.3064 & 0.6180 & 0.2215 & 0.4654 & 0.7649 & 1.0844 & 0.0615 & 0.3643 & 0.1946 \\
      Ru & 0.3060 & 0.6201 & 0.2104 & 0.4668 & 0.7761 & 1.0800 & 0.0663 & 0.3421 & 0.2026 \\
      Sb & 0.3050 & 0.6244 & 0.1461 & 0.4696 & 0.7473 & 1.1052 & 0.0689 & 0.3676 & 0.2043 \\
      Sn & 0.3115 & 0.5941 & 0.1441 & 0.4373 & 0.7202 & 1.0761 & 0.0591 & 0.3397 & 0.1601 \\
      Tc & 0.3170 & 0.6102 & 0.2178 & 0.4543 & 0.7890 & 1.0807 & 0.0698 & 0.3442 & 0.2008 \\
      Y  & 0.2768 & 0.5949 & 0.1712 & 0.4467 & 0.7278 & 1.0780 & 0.0573 & 0.3206 & 0.1349 \\
      Zr & 0.2939 & 0.6068 & 0.1888 & 0.4631 & 0.7737 & 1.0875 & 0.0744 & 0.3155 & 0.1592 \\
      Au & 0.3123 & 0.6245 & 0.2147 & 0.4761 & 0.7312 & 1.0845 & 0.0527 & 0.3432 & 0.2066 \\
      Bi & 0.3076 & 0.6148 & 0.1485 & 0.4612 & 0.7392 & 1.0935 & 0.0611 & 0.3554 & 0.1893 \\
      Hf & 0.2928 & 0.6104 & 0.1880 & 0.4658 & 0.7769 & 1.0954 & 0.0765 & 0.3218 & 0.1728 \\
      Hg & 0.3220 & 0.6221 & 0.1486 & 0.4707 & 0.7338 & 1.0809 & 0.0541 & 0.3469 & 0.1924 \\
      Ir & 0.3127 & 0.6311 & 0.2271 & 0.4818 & 0.7530 & 1.0740 & 0.1084 & 0.3314 & 0.2142 \\
      Os & 0.3122 & 0.6260 & 0.2258 & 0.4796 & 0.7531 & 1.0911 & 0.0670 & 0.3501 & 0.2154 \\
      Pb & 0.3152 & 0.5834 & 0.1401 & 0.4239 & 0.7113 & 1.0486 & 0.0513 & 0.3043 & 0.1352 \\
      Pt & 0.3144 & 0.6361 & 0.2225 & 0.4908 & 0.7569 & 1.0921 & 0.0603 & 0.3672 & 0.2058 \\
      Re & 0.3171 & 0.6175 & 0.2213 & 0.4665 & 0.8021 & 1.0985 & 0.0780 & 0.3597 & 0.2032 \\
      Ta & 0.3099 & 0.5977 & 0.2000 & 0.4414 & 0.7993 & 1.0880 & 0.0847 & 0.3410 & 0.1780 \\
      Tl & 0.3114 & 0.5614 & 0.1389 & 0.3494 & 0.7105 & 1.0410 & 0.0246 & 0.3210 & 0.0951 \\
      W  & 0.3146 & 0.6059 & 0.2160 & 0.4456 & 0.8033 & 1.0886 & 0.0725 & 0.3387 & 0.1834 \\
      \bottomrule
    \end{tabular}
  }
\end{table}

% Supplementary Table 8: Entropy corrections for adsorbate@g-C3N4
\begin{table}[htbp]
\label{supp_table8:s_g-c3n4}
  \caption{Entropy corrections ($-T \cdot S$) in eV for relaxed
    adsorbates supported on g-C\textsubscript{3}N\textsubscript{4} at 298.15 K.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{l *{9}{l}}
      \toprule
      Metal & *CO\textsubscript{2} & *COOH& *CO     & *CHO
      & *CH\textsubscript{2}O& *OCH\textsubscript{3}& *O      & *OH     & *H      \\
      \midrule
      Al & -0.2195 & -0.2476 & -0.1874 & -0.1939 & -0.2359 & -0.1759 & -0.0894 & -0.1010 & -0.0124 \\
      Co & -0.2359 & -0.2665 & -0.1544 & -0.1572 & -0.1755 & -0.2095 & -0.1078 & -0.1510 & -0.0341 \\
      Cr & -0.1862 & -0.2605 & -0.2075 & -0.2069 & -0.1478 & -0.1733 & -0.0790 & -0.1437 & -0.0221 \\
      Cu & -0.3098 & -0.2403 & -0.1655 & -0.1803 & -0.1647 & -0.2378 & -0.0977 & -0.1375 & -0.0297 \\
      Fe & -0.2184 & -0.2028 & -0.1676 & -0.1797 & -0.1742 & -0.2640 & -0.0815 & -0.1285 & -0.0215 \\
      Ga & -0.2602 & -0.2558 & -0.2291 & -0.1881 & -0.1252 & -0.2576 & -0.0897 & -0.1441 & -0.0182 \\
      Ge & -0.1978 & -0.2423 & -0.1635 & -0.1802 & -0.2104 & -0.2364 & -0.1178 & -0.1052 & -0.0087 \\
      Mn & -0.2142 & -0.2678 & -0.1863 & -0.2059 & -0.1788 & -0.2752 & -0.0808 & -0.1425 & -0.0261 \\
      Ni & -0.2460 & -0.1948 & -0.1592 & -0.1224 & -0.1275 & -0.2506 & -0.0997 & -0.1384 & -0.0224 \\
      Sc & -0.2227 & -0.2369 & -0.1843 & -0.1667 & -0.1817 & -0.2035 & -0.0728 & -0.1416 & -0.0226 \\
      Ti & -0.1978 & -0.2263 & -0.1685 & -0.1389 & -0.1414 & -0.1795 & -0.0578 & -0.1013 & -0.0195 \\
      V  & -0.2603 & -0.2511 & -0.1452 & -0.2028 & -0.1360 & -0.2374 & -0.0612 & -0.1082 & -0.0186 \\
      Zn & -0.2928 & -0.2543 & -0.1909 & -0.1959 & -0.2202 & -0.2331 & -0.0910 & -0.1185 & -0.0228 \\
      Ag & -0.2458 & -0.2428 & -0.1910 & -0.1890 & -0.3097 & -0.2421 & -0.1060 & -0.1392 & -0.0276 \\
      Cd & -0.2961 & -0.2561 & -0.2368 & -0.2023 & -0.2076 & -0.2630 & -0.0270 & -0.1408 & -0.0337 \\
      In & -0.1978 & -0.2895 & -0.1071 & -0.1840 & -0.1338 & -0.2826 & -0.1011 & -0.1386 & -0.0405 \\
      Mo & -0.1912 & -0.2310 & -0.1310 & -0.1680 & -0.1269 & -0.2228 & -0.0641 & -0.1496 & -0.0155 \\
      Nb & -0.1896 & -0.2540 & -0.1460 & -0.1924 & -0.1290 & -0.2147 & -0.0575 & -0.1177 & -0.0261 \\
      Pd & -0.2405 & -0.2385 & -0.1649 & -0.1740 & -0.2538 & -0.2201 & -0.0900 & -0.1143 & -0.0518 \\
      Rh & -0.2408 & -0.2385 & -0.1355 & -0.1734 & -0.1853 & -0.2247 & -0.0863 & -0.0866 & -0.0222 \\
      Ru & -0.2197 & -0.2304 & -0.1450 & -0.1769 & -0.1655 & -0.1574 & -0.0832 & -0.1294 & -0.0138 \\
      Sb & -0.2354 & -0.2277 & -0.1320 & -0.1793 & -0.2211 & -0.1961 & -0.0616 & -0.0808 & -0.0090 \\
      Sn & -0.2606 & -0.2544 & -0.2227 & -0.2021 & -0.3045 & -0.2568 & -0.0787 & -0.1171 & -0.0160 \\
      Tc & -0.1875 & -0.2373 & -0.1347 & -0.1733 & -0.1485 & -0.2301 & -0.0697 & -0.1007 & -0.0121 \\
      Y  & -0.2557 & -0.2528 & -0.1929 & -0.1886 & -0.2130 & -0.2081 & -0.0824 & -0.1393 & -0.0285 \\
      Zr & -0.2083 & -0.2325 & -0.1624 & -0.1459 & -0.1574 & -0.2014 & -0.0640 & -0.1556 & -0.0208 \\
      Au & -0.1901 & -0.2267 & -0.1484 & -0.1846 & -0.2939 & -0.2307 & -0.0857 & -0.1271 & -0.0138 \\
      Bi & -0.2366 & -0.2392 & -0.1915 & -0.1899 & -0.2361 & -0.1986 & -0.0683 & -0.0934 & -0.0100 \\
      Hf & -0.2043 & -0.2219 & -0.1618 & -0.1418 & -0.1489 & -0.1981 & -0.0610 & -0.1414 & -0.0163 \\
      Hg & -0.2980 & -0.2371 & -0.2812 & -0.1995 & -0.3039 & -0.2465 & -0.0878 & -0.1245 & -0.0175 \\
      Ir & -0.2097 & -0.2290 & -0.1261 & -0.1803 & -0.1913 & -0.2221 & -0.0377 & -0.1379 & -0.0159 \\
      Os & -0.1795 & -0.2235 & -0.1294 & -0.1519 & -0.1609 & -0.2027 & -0.0680 & -0.1093 & -0.0124 \\
      Pb & -0.2546 & -0.2691 & -0.1666 & -0.2165 & -0.2641 & -0.2520 & -0.1007 & -0.1097 & -0.0257 \\
      Pt & -0.2411 & -0.2029 & -0.1496 & -0.1593 & -0.2009 & -0.2246 & -0.0950 & -0.0937 & -0.0137 \\
      Re & -0.1869 & -0.2262 & -0.1293 & -0.1697 & -0.1266 & -0.2165 & -0.0644 & -0.0953 & -0.0146 \\
      Ta & -0.1872 & -0.2532 & -0.1523 & -0.1847 & -0.1265 & -0.1798 & -0.0535 & -0.1040 & -0.0177 \\
      Tl & -0.2599 & -0.2833 & -0.1807 & -0.1538 & -0.2422 & -0.1501 & -0.0126 & -0.1338 & -0.0533 \\
      W  & -0.1791 & -0.2368 & -0.1335 & -0.1723 & -0.1225 & -0.2253 & -0.0673 & -0.1180 & -0.0201 \\
      \bottomrule
    \end{tabular}
  }
\end{table}

% Supplementary Table 9: ZPE for adsorbates supported on N-doped graphene
\begin{table}[htbp]
\label{supp_table9:zpe_n_gra}
  \caption{Zero-point energies in eV for relaxed adsorbates
    supported on nitrogen-doped graphene at 298.15 K.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{l *{9}{c}}
      \toprule
      Metal & *CO\textsubscript{2} & *COOH & *CO & *CHO
      & *CH\textsubscript{2}O & *OCH\textsubscript{3} & *O & *OH    & *H     \\
      \midrule
      Al & 0.3093 & 0.6024 & 0.1683 & 0.4395 & 0.7509 & 1.0855 & 0.0552 & 0.3401 & 0.1864 \\
      Co & 0.3140 & 0.6247 & 0.2044 & 0.4787 & 0.7345 & 1.0615 & 0.0606 & 0.3315 & 0.2006 \\
      Cr & 0.3108 & 0.5983 & 0.2007 & 0.4410 & 0.7211 & 1.0824 & 0.0828 & 0.3253 & 0.1625 \\
      Cu & 0.3145 & 0.5943 & 0.1460 & 0.4377 & 0.7196 & 1.0160 & 0.0407 & 0.3104 & 0.1523 \\
      Fe & 0.3117 & 0.6148 & 0.2195 & 0.4634 & 0.7877 & 1.0678 & 0.0703 & 0.3163 & 0.1858 \\
      Ga & 0.3129 & 0.6080 & 0.1561 & 0.4537 & 0.7495 & 1.0826 & 0.0573 & 0.3420 & 0.1897 \\
      Ge & 0.3131 & 0.6206 & 0.1402 & 0.4629 & 0.7281 & 1.1014 & 0.0686 & 0.3538 & 0.2110 \\
      Mn & 0.3120 & 0.6111 & 0.2128 & 0.4601 & 0.8078 & 1.0640 & 0.0793 & 0.3148 & 0.2040 \\
      Ni & 0.3110 & 0.6059 & 0.1485 & 0.4602 & 0.7283 & 1.0467 & 0.0402 & 0.3168 & 0.1620 \\
      Sc & 0.2741 & 0.5910 & 0.1620 & 0.4456 & 0.7019 & 1.0758 & 0.0575 & 0.3166 & 0.1338 \\
      Ti & 0.2920 & 0.6020 & 0.1781 & 0.4525 & 0.7593 & 1.0831 & 0.0731 & 0.2981 & 0.1481 \\
      V  & 0.2984 & 0.6055 & 0.1865 & 0.4609 & 0.7756 & 1.0843 & 0.0786 & 0.3171 & 0.1525 \\
      Zn & 0.3143 & 0.5928 & 0.1554 & 0.4357 & 0.7116 & 1.0521 & 0.0394 & 0.3248 & 0.1640 \\
      Ag & 0.3127 & 0.5862 & 0.1862 & 0.4307 & 0.7518 & 1.0419 & 0.0356 & 0.3231 & 0.1308 \\
      Cd & 0.3055 & 0.5910 & 0.1518 & 0.4402 & 0.7312 & 1.0524 & 0.0366 & 0.3189 & 0.1483 \\
      In & 0.3096 & 0.5986 & 0.1414 & 0.4477 & 0.7152 & 1.0724 & 0.0437 & 0.3300 & 0.1609 \\
      Mo & 0.2949 & 0.6111 & 0.1914 & 0.4728 & 0.7806 & 1.0887 & 0.0795 & 0.3253 & 0.1500 \\
      Nb & 0.2946 & 0.5996 & 0.1767 & 0.4601 & 0.7874 & 1.0896 & 0.0742 & 0.3048 & 0.1515 \\
      Pd & 0.3122 & 0.5993 & 0.1444 & 0.4490 & 0.7303 & 1.0342 & 0.0300 & 0.3047 & 0.1558 \\
      Rh & 0.3115 & 0.6288 & 0.1969 & 0.4826 & 0.7388 & 1.0867 & 0.0602 & 0.3453 & 0.2025 \\
      Ru & 0.3152 & 0.6122 & 0.2180 & 0.4569 & 0.7918 & 1.0947 & 0.0698 & 0.3483 & 0.1873 \\
      Sb & 0.3115 & 0.5874 & 0.1397 & 0.4349 & 0.7184 & 1.0467 & 0.0646 & 0.3351 & 0.1752 \\
      Sn & 0.3112 & 0.6002 & 0.1400 & 0.4367 & 0.7239 & 1.0836 & 0.0568 & 0.3359 & 0.1710 \\
      Tc & 0.3053 & 0.6110 & 0.2105 & 0.4251 & 0.7934 & 1.0870 & 0.0807 & 0.3496 & 0.1741 \\
      Y  & 0.2709 & 0.5835 & 0.1580 & 0.4307 & 0.7059 & 1.0703 & 0.0500 & 0.3200 & 0.1226 \\
      Zr & 0.2719 & 0.5963 & 0.1723 & 0.4430 & 0.7327 & 1.0846 & 0.0669 & 0.3176 & 0.1472 \\
      Au & 0.3137 & 0.5480 & 0.1442 & 0.3577 & 0.7209 & 0.9864 & 0.0174 & 0.2881 & 0.1027 \\
      Bi & 0.3111 & 0.5778 & 0.1378 & 0.4190 & 0.7133 & 1.0475 & 0.0550 & 0.3241 & 0.1533 \\
      Hf & 0.2699 & 0.6002 & 0.1714 & 0.4452 & 0.7372 & 1.0904 & 0.0687 & 0.3242 & 0.1569 \\
      Hg & 0.3113 & 0.6082 & 0.1397 & 0.4569 & 0.7258 & 1.0720 & 0.0412 & 0.3304 & 0.1625 \\
      Ir & 0.3132 & 0.6309 & 0.2106 & 0.4856 & 0.7390 & 1.0888 & 0.0626 & 0.3456 & 0.2022 \\
      Os & 0.3156 & 0.6148 & 0.2179 & 0.4592 & 0.7971 & 1.0984 & 0.0768 & 0.3463 & 0.1812 \\
      Pb & 0.3116 & 0.5352 & 0.1385 & 0.3677 & 0.7147 & 1.0532 & 0.0474 & 0.3172 & 0.1302 \\
      Pt & 0.3139 & 0.6134 & 0.1439 & 0.4671 & 0.7359 & 1.0406 & 0.0482 & 0.3114 & 0.1736 \\
      Re & 0.3035 & 0.6045 & 0.2095 & 0.4566 & 0.7995 & 1.0957 & 0.0846 & 0.3493 & 0.1818 \\
      Ta & 0.2751 & 0.6019 & 0.1869 & 0.4630 & 0.7724 & 1.0937 & 0.0751 & 0.3060 & 0.1669 \\
      Tl & 0.3131 & 0.5988 & 0.1402 & 0.4439 & 0.7193 & 1.0169 & 0.0573 & 0.3244 & 0.1506 \\
      W  & 0.2889 & 0.6117 & 0.1927 & 0.4777 & 0.7852 & 1.0941 & 0.0810 & 0.3273 & 0.1832 \\
      \bottomrule
    \end{tabular}
  }
\end{table}

% Supplementary Table 10: Entropy corrections for adsorbate@N-doped graphene
\begin{table}[htbp]
\label{supp_table10:s_n_gra}
  \caption{Entropy corrections ($-T \cdot S$) in eV for relaxed
    adsorbates supported on nitrogen-doped graphene at 298.15 K.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{l *{9}{l}}
      \toprule
      Metal & *CO\textsubscript{2} & *COOH& *CO     & *CHO
      & *CH\textsubscript{2}O& *OCH\textsubscript{3}& *O      & *OH     & *H      \\
      \midrule
      Al & -0.1981 & -0.2564 & -0.1979 & -0.1424 & -0.2051 & -0.1839 & -0.0794 & -0.1211 & -0.0113 \\
      Co & -0.2358 & -0.2213 & -0.1589 & -0.1584 & -0.2025 & -0.2223 & -0.0701 & -0.0817 & -0.0123 \\
      Cr & -0.2593 & -0.1973 & -0.1568 & -0.2037 & -0.2291 & -0.1679 & -0.0569 & -0.0875 & -0.0228 \\
      Cu & -0.2457 & -0.2636 & -0.2132 & -0.2040 & -0.1267 & -0.2137 & -0.1188 & -0.1711 & -0.0226 \\
      Fe & -0.1936 & -0.2385 & -0.1356 & -0.1758 & -0.1985 & -0.1910 & -0.0647 & -0.0892 & -0.0172 \\
      Ga & -0.2565 & -0.1937 & -0.1600 & -0.2025 & -0.2139 & -0.1766 & -0.0727 & -0.1269 & -0.0126 \\
      Ge & -0.2469 & -0.2413 & -0.1027 & -0.1937 & -0.1687 & -0.1551 & -0.0690 & -0.1091 & -0.0080 \\
      Mn & -0.1929 & -0.2402 & -0.1409 & -0.1736 & -0.1949 & -0.2011 & -0.0563 & -0.1631 & -0.0160 \\
      Ni & -0.1239 & -0.2440 & -0.2027 & -0.1785 & -0.2318 & -0.1792 & -0.0929 & -0.1520 & -0.0170 \\
      Sc & -0.1938 & -0.2647 & -0.1418 & -0.2024 & -0.2291 & -0.2139 & -0.0790 & -0.1516 & -0.0293 \\
      Ti & -0.2093 & -0.2423 & -0.1798 & -0.1640 & -0.1765 & -0.1433 & -0.0692 & -0.1462 & -0.0262 \\
      V  & -0.2112 & -0.2373 & -0.1734 & -0.1578 & -0.1642 & -0.0767 & -0.0621 & -0.1522 & -0.0273 \\
      Zn & -0.2567 & -0.2001 & -0.2491 & -0.2135 & -0.1581 & -0.1396 & -0.1091 & -0.1668 & -0.0211 \\
      Ag & -0.2600 & -0.2035 & -0.1802 & -0.2231 & -0.1723 & -0.1531 & -0.1293 & -0.1487 & -0.0596 \\
      Cd & -0.2615 & -0.2042 & -0.1827 & -0.2185 & -0.2727 & -0.2094 & -0.1192 & -0.1850 & -0.0320 \\
      In & -0.1886 & -0.2021 & -0.1593 & -0.2221 & -0.1949 & -0.2061 & -0.1158 & -0.1603 & -0.0238 \\
      Mo & -0.2049 & -0.2075 & -0.1659 & -0.1327 & -0.1650 & -0.2041 & -0.0605 & -0.1258 & -0.0437 \\
      Nb & -0.2072 & -0.2300 & -0.1145 & -0.1439 & -0.1403 & -0.2086 & -0.0684 & -0.1863 & -0.0251 \\
      Pd & -0.1843 & -0.2620 & -0.1476 & -0.2000 & -0.1645 & -0.2057 & -0.1131 & -0.1688 & -0.0223 \\
      Rh & -0.2986 & -0.2220 & -0.1669 & -0.1667 & -0.1953 & -0.1423 & -0.0693 & -0.1229 & -0.0149 \\
      Ru & -0.1998 & -0.1779 & -0.1375 & -0.1827 & -0.1455 & -0.2017 & -0.0653 & -0.1103 & -0.0202 \\
      Sb & -0.1938 & -0.1984 & -0.1084 & -0.2024 & -0.1970 & -0.0996 & -0.0780 & -0.1183 & -0.0109 \\
      Sn & -0.1931 & -0.2040 & -0.1708 & -0.2214 & -0.1801 & -0.2006 & -0.0911 & -0.1474 & -0.0185 \\
      Tc & -0.2038 & -0.2277 & -0.1483 & -0.1505 & -0.1515 & -0.1810 & -0.0569 & -0.1008 & -0.0234 \\
      Y  & -0.1994 & -0.2704 & -0.1507 & -0.1605 & -0.2919 & -0.2162 & -0.0981 & -0.1548 & -0.0344 \\
      Zr & -0.2233 & -0.2608 & -0.1845 & -0.1849 & -0.2016 & -0.2144 & -0.0776 & -0.1500 & -0.0242 \\
      Au & -0.1872 & -0.1923 & -0.2237 & -0.1738 & -0.1907 & -0.2639 & -0.1497 & -0.1276 & -0.0301 \\
      Bi & -0.1978 & -0.2736 & -0.1731 & -0.1581 & -0.2369 & -0.0995 & -0.0937 & -0.1386 & -0.0171 \\
      Hf & -0.2235 & -0.2562 & -0.1836 & -0.1839 & -0.1987 & -0.2082 & -0.0759 & -0.1417 & -0.0211 \\
      Hg & -0.1907 & -0.1929 & -0.1053 & -0.0955 & -0.1765 & -0.2558 & -0.1157 & -0.1694 & -0.0357 \\
      Ir & -0.2967 & -0.2218 & -0.1486 & -0.1659 & -0.2038 & -0.2134 & -0.0723 & -0.1326 & -0.0200 \\
      Os & -0.1948 & -0.2462 & -0.1422 & -0.1945 & -0.1448 & -0.2046 & -0.0598 & -0.1205 & -0.0322 \\
      Pb & -0.1979 & -0.2320 & -0.1108 & -0.2414 & -0.1437 & -0.2639 & -0.1126 & -0.1465 & -0.0238 \\
      Pt & -0.1770 & -0.2479 & -0.1481 & -0.1915 & -0.2820 & -0.2030 & -0.0791 & -0.1666 & -0.0228 \\
      Re & -0.1958 & -0.2535 & -0.1508 & -0.1867 & -0.1462 & -0.2483 & -0.0561 & -0.1018 & -0.0249 \\
      Ta & -0.2078 & -0.2197 & -0.1675 & -0.1373 & -0.1542 & -0.2084 & -0.0698 & -0.1979 & -0.0195 \\
      Tl & -0.2573 & -0.2043 & -0.1719 & -0.2270 & -0.1288 & -0.2326 & -0.0688 & -0.1876 & -0.0332 \\
      W  & -0.2007 & -0.2007 & -0.1642 & -0.1194 & -0.1565 & -0.2032 & -0.0607 & -0.1275 & -0.0202 \\
      \bottomrule
    \end{tabular}
  }
\end{table}

% Supplementary Table 11: ZPE for adsorbates supported on graphene with dual-vac
\begin{table}[htbp]
\label{supp_table11:zpe_vac_gra}
  \caption{Zero-point energies in eV for relaxed adsorbates
    supported on graphene with dual-vacancy at 298.15 K.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{l *{9}{l}}
      \toprule
      Metal & *CO\textsubscript{2} & *COOH & *CO & *CHO
      & *CH\textsubscript{2}O & *OCH\textsubscript{3}          & *O     & *OH    & *H     \\
      \midrule
      Al & 0.3140 & 0.5988 & 0.1888 & 0.4386 & 0.7218 & 1.0769 & 0.0472 & 0.3366 & 0.1729 \\
      Co & 0.3080 & 0.6179 & 0.2173 & 0.4899 & 0.7928 & 1.0616 & 0.0573 & 0.3458 & 0.1726 \\
      Cr & 0.3318 & 0.5862 & 0.1957 & 0.4277 & 0.8826 & 1.0733 & 0.0792 & 0.3294 & 0.1560 \\
      Cu & 0.3140 & 0.6023 & 0.1769 & 0.5229 & 0.7172 & 1.0131 & 0.0323 & 0.3298 & 0.2857 \\
      Fe & 0.3177 & 0.6110 & 0.2083 & 0.5038 & 0.7828 & 1.0802 & 0.0776 & 0.3515 & 0.1987 \\
      Ga & 0.3126 & 0.6056 & 0.1492 & 0.4494 & 0.7196 & 1.0788 & 0.0446 & 0.3370 & 0.1810 \\
      Ge & 0.3138 & 0.6171 & 0.1471 & 0.4657 & 0.7196 & 1.0935 & 0.0540 & 0.3514 & 0.2028 \\
      Mn & 0.3167 & 0.6141 & 0.2050 & 0.4803 & 0.8009 & 1.0809 & 0.0756 & 0.3411 & 0.1594 \\
      Ni & 0.3134 & 0.6251 & 0.1942 & 0.4571 & 0.7230 & 1.0738 & 0.0496 & 0.3521 & 0.2877 \\
      Sc & 0.3229 & 0.5874 & 0.1668 & 0.4367 & 0.7358 & 1.0673 & 0.0533 & 0.3129 & 0.1288 \\
      Ti & 0.3240 & 0.6049 & 0.1749 & 0.4562 & 0.7202 & 1.0694 & 0.0670 & 0.3143 & 0.1559 \\
      V  & 0.2937 & 0.6042 & 0.1933 & 0.4593 & 0.8437 & 1.0745 & 0.0753 & 0.3319 & 0.1687 \\
      Zn & 0.3123 & 0.5888 & 0.1564 & 0.4317 & 0.7111 & 1.0363 & 0.0324 & 0.3157 & 0.1363 \\
      Ag & 0.3172 & 0.5984 & 0.1399 & 0.4312 & 0.7199 & 1.0034 & 0.0258 & 0.3141 & 0.2932 \\
      Cd & 0.3120 & 0.5796 & 0.1487 & 0.4176 & 0.7144 & 1.0198 & 0.0278 & 0.3084 & 0.1280 \\
      In & 0.3134 & 0.5973 & 0.1458 & 0.4443 & 0.7170 & 1.0666 & 0.0418 & 0.3236 & 0.1610 \\
      Mo & 0.3052 & 0.5923 & 0.1942 & 0.4668 & 0.7718 & 1.0844 & 0.0745 & 0.3350 & 0.1755 \\
      Nb & 0.2952 & 0.6071 & 0.1862 & 0.4605 & 0.7505 & 1.0749 & 0.0721 & 0.3274 & 0.1717 \\
      Pd & 0.3151 & 0.6191 & 0.1872 & 0.4609 & 0.7187 & 1.0525 & 0.0397 & 0.3371 & 0.2963 \\
      Rh & 0.2870 & 0.6096 & 0.2076 & 0.4297 & 0.7850 & 1.0628 & 0.0604 & 0.3590 & 0.2678 \\
      Ru & 0.3000 & 0.6040 & 0.2097 & 0.4645 & 0.7812 & 1.0710 & 0.0790 & 0.3319 & 0.1758 \\
      Sb & 0.3122 & 0.6143 & 0.1436 & 0.4604 & 0.7258 & 1.0938 & 0.0608 & 0.3481 & 0.1912 \\
      Sn & 0.3132 & 0.6077 & 0.1493 & 0.4550 & 0.7182 & 1.0807 & 0.0517 & 0.3338 & 0.1770 \\
      Tc & 0.3051 & 0.6025 & 0.2086 & 0.4585 & 0.7777 & 1.0690 & 0.0799 & 0.3285 & 0.2898 \\
      Y  & 0.3219 & 0.5847 & 0.1632 & 0.4299 & 0.7245 & 1.0646 & 0.0557 & 0.3182 & 0.1164 \\
      Zr & 0.3213 & 0.5966 & 0.1669 & 0.4527 & 0.7200 & 1.0620 & 0.0568 & 0.3151 & 0.1510 \\
      Au & 0.3172 & 0.5927 & 0.1448 & 0.4677 & 0.7261 & 0.9827 & 0.0291 & 0.2904 & 0.2898 \\
      Bi & 0.3108 & 0.6062 & 0.1406 & 0.4521 & 0.7185 & 1.0309 & 0.0528 & 0.3289 & 0.1696 \\
      Hf & 0.3241 & 0.5981 & 0.1713 & 0.4543 & 0.7236 & 1.0751 & 0.0586 & 0.3142 & 0.1550 \\
      Hg & 0.3130 & 0.5763 & 0.1421 & 0.3959 & 0.7161 & 0.9951 & 0.0177 & 0.2917 & 0.1258 \\
      Ir & 0.3054 & 0.6129 & 0.2147 & 0.4199 & 0.7939 & 1.0680 & 0.0712 & 0.3603 & 0.1863 \\
      Os & 0.3033 & 0.6068 & 0.2155 & 0.4695 & 0.7674 & 1.0779 & 0.0823 & 0.3330 & 0.1401 \\
      Pb & 0.3118 & 0.6044 & 0.1424 & 0.4518 & 0.7199 & 1.0740 & 0.0502 & 0.3250 & 0.1623 \\
      Pt & 0.3124 & 0.6242 & 0.1985 & 0.4680 & 0.7263 & 1.0514 & 0.0477 & 0.3406 & 0.2725 \\
      Re & 0.3033 & 0.6034 & 0.2108 & 0.4601 & 0.7864 & 1.0807 & 0.0780 & 0.3317 & 0.2191 \\
      Ta & 0.2938 & 0.6078 & 0.1872 & 0.4601 & 0.7707 & 1.0754 & 0.0732 & 0.3243 & 0.1783 \\
      Tl & 0.3091 & 0.5948 & 0.1424 & 0.4406 & 0.7123 & 1.0568 & 0.0342 & 0.3173 & 0.1569 \\
      W  & 0.3066 & 0.6062 & 0.1912 & 0.3984 & 0.8010 & 1.0777 & 0.0787 & 0.3333 & 0.1815 \\
      \bottomrule
    \end{tabular}
  }
\end{table}

% Supplementary Table 12: Entropy corrections for adsorbate@graphene with dual-vac
\begin{table}[htbp]
\label{supp_table12:s_vac_gra}
  \caption{Entropy corrections ($-T \cdot S$) in eV for relaxed
    adsorbates supported on graphene with dual-vacancy at 298.15 K.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{l *{9}{l}}
      \toprule
        Metal & *CO\textsubscript{2} & *COOH& *CO     & *CHO
        & *CH\textsubscript{2}O& *OCH\textsubscript{3}& *O      & *OH     & *H      \\
        \midrule
        Al & -0.2501 & -0.2622 & -0.1799 & -0.2106 & -0.3101 & -0.2050 & -0.0980 & -0.1303 & -0.0165 \\
        Co & -0.1962 & -0.2352 & -0.1529 & -0.1520 & -0.1515 & -0.0821 & -0.0427 & -0.1165 & -0.0376 \\
        Cr & -0.2178 & -0.2281 & -0.1603 & -0.2297 & -0.0849 & -0.0961 & -0.0623 & -0.1200 & -0.0250 \\
        Cu & -0.1907 & -0.2528 & -0.1276 & -0.1239 & -0.1974 & -0.2279 & -0.0711 & -0.1357 & -0.0044 \\
        Fe & -0.2328 & -0.2473 & -0.0848 & -0.1300 & -0.1611 & -0.1248 & -0.0629 & -0.0996 & -0.0292 \\
        Ga & -0.1842 & -0.2577 & -0.2003 & -0.2061 & -0.1912 & -0.2332 & -0.0956 & -0.1395 & -0.0145 \\
        Ge & -0.2491 & -0.1850 & -0.1442 & -0.1942 & -0.1974 & -0.1522 & -0.0886 & -0.1119 & -0.0101 \\
        Mn & -0.2384 & -0.2493 & -0.1559 & -0.1654 & -0.1993 & -0.1304 & -0.0599 & -0.1090 & -0.0275 \\
        Ni & -0.2497 & -0.2279 & -0.1683 & -0.1836 & -0.2402 & -0.2254 & -0.1159 & -0.1027 & -0.0021 \\
        Sc & -0.2252 & -0.1996 & -0.1449 & -0.2023 & -0.2586 & -0.2259 & -0.1128 & -0.1670 & -0.0322 \\
        Ti & -0.2137 & -0.2297 & -0.1918 & -0.1703 & -0.2184 & -0.2455 & -0.0745 & -0.1598 & -0.0186 \\
        V  & -0.2319 & -0.2441 & -0.1693 & -0.1713 & -0.0933 & -0.0952 & -0.0673 & -0.1171 & -0.0164 \\
        Zn & -0.1971 & -0.2071 & -0.1798 & -0.2147 & -0.1577 & -0.2074 & -0.1203 & -0.1637 & -0.0293 \\
        Ag & -0.2465 & -0.2579 & -0.1696 & -0.1739 & -0.1930 & -0.2345 & -0.0117 & -0.1000 & -0.0038 \\
        Cd & -0.1963 & -0.2145 & -0.2124 & -0.2350 & -0.2120 & -0.2341 & -0.1338 & -0.1754 & -0.0354 \\
        In & -0.2567 & -0.2013 & -0.1464 & -0.1558 & -0.2003 & -0.1995 & -0.1043 & -0.1065 & -0.0228 \\
        Mo & -0.2106 & -0.2086 & -0.1656 & -0.1746 & -0.1628 & -0.2038 & -0.0668 & -0.1129 & -0.0144 \\
        Nb & -0.2307 & -0.2394 & -0.1752 & -0.1660 & -0.1679 & -0.1668 & -0.0688 & -0.1288 & -0.0138 \\
        Pd & -0.3134 & -0.2362 & -0.1751 & -0.1707 & -0.1779 & -0.0819 & -0.0677 & -0.1222 & -0.0039 \\
        Rh & -0.2128 & -0.2468 & -0.1542 & -0.1986 & -0.1595 & -0.0720 & -0.0949 & -0.0933 & -0.0021 \\
        Ru & -0.2275 & -0.2613 & -0.1426 & -0.1760 & -0.1693 & -0.0786 & -0.0577 & -0.1332 & -0.0175 \\
        Sb & -0.1891 & -0.1940 & -0.2266 & -0.1526 & -0.2398 & -0.2284 & -0.0788 & -0.1157 & -0.0132 \\
        Sn & -0.2529 & -0.1968 & -0.2028 & -0.2165 & -0.1898 & -0.1845 & -0.0886 & -0.1568 & -0.0170 \\
        Tc & -0.2158 & -0.2603 & -0.1459 & -0.1731 & -0.1578 & -0.0973 & -0.0574 & -0.1222 & -0.0647 \\
        Y  & -0.2962 & -0.2657 & -0.1496 & -0.2075 & -0.2598 & -0.2231 & -0.1096 & -0.1662 & -0.0384 \\
        Zr & -0.1548 & -0.2417 & -0.1360 & -0.1770 & -0.2745 & -0.2015 & -0.0919 & -0.1551 & -0.0185 \\
        Au & -0.2413 & -0.2076 & -0.2877 & -0.1138 & -0.2448 & -0.2634 & -0.0095 & -0.1447 & -0.0041 \\
        Bi & -0.1970 & -0.2006 & -0.2337 & -0.1648 & -0.1315 & -0.2394 & -0.0927 & -0.1399 & -0.0183 \\
        Hf & -0.2728 & -0.2398 & -0.1920 & -0.1784 & -0.2711 & -0.3070 & -0.0834 & -0.1620 & -0.0186 \\
        Hg & -0.1953 & -0.2848 & -0.1710 & -0.1987 & -0.2065 & -0.1285 & -0.1491 & -0.1469 & -0.0370 \\
        Ir & -0.2144 & -0.1796 & -0.1476 & -0.1609 & -0.1474 & -0.0751 & -0.0699 & -0.0953 & -0.0258 \\
        Os & -0.2163 & -0.2409 & -0.1420 & -0.1599 & -0.1776 & -0.0816 & -0.0565 & -0.1218 & -0.0724 \\
        Pb & -0.1964 & -0.2001 & -0.2339 & -0.1644 & -0.1894 & -0.1890 & -0.0888 & -0.1086 & -0.0227 \\
        Pt & -0.1825 & -0.2340 & -0.1657 & -0.1759 & -0.2307 & -0.1921 & -0.0612 & -0.1216 & -0.0022 \\
        Re & -0.2117 & -0.2566 & -0.1472 & -0.1661 & -0.1530 & -0.1479 & -0.0619 & -0.1223 & -0.0261 \\
        Ta & -0.2269 & -0.2462 & -0.1744 & -0.1715 & -0.1696 & -0.1787 & -0.0694 & -0.1428 & -0.0135 \\
        Tl & -0.1981 & -0.2044 & -0.2304 & -0.2198 & -0.2048 & -0.2595 & -0.1146 & -0.1764 & -0.0256 \\
        W  & -0.2006 & -0.2433 & -0.1670 & -0.1453 & -0.1295 & -0.1025 & -0.0613 & -0.1185 & -0.0157 \\
        \bottomrule
    \end{tabular}
  }
\end{table}

\subsection{Correlation analysis of adsorption energies}
\label{supp_sec2.4_corr_analysis}

\cref{supp_fig9:corr_eads} presents a Pearson correlation map that illustrates the relationships between the adsorption energies of various intermediates. This analysis encompasses all six substrates under consideration. The mean Pearson correlation coefficient is 0.6023 with a variance of 0.0109, signifying a robust correlation between the adsorption energies of different intermediates, accompanied by low variability.

% SI Figure 9: Pearson correlation map for Eads of adsorbates for all substrates
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7\textwidth]{supp_fig9_corr_eads.pdf}
  \caption{\textbf{Pearson correlation map for adsorption energies of various adsorbates for all six substrates.}}
  \label{supp_fig9:corr_eads}
\end{figure}

\subsection{The scaling relation scheme and the hybrid scaling relation}
\label{supp_sec2.5_scaling}

In line with the scaling relation framework proposed by Abild-Pedersen et al. \cite{abild2007scaling},
and later applied to the CO\textsubscript{2} reduction to methane process by Peterson \& Nørskov \cite{peterson2012activity},
adsorbates in the CO\textsubscript{2}RR process can be categorized into C-centered or O-centered species.
In each group, a representative species is designated as the ``descriptor'', allowing approximating the adsorption energies of other species within the same group using scaling relations.
The following equations capture this relationship:

\begin{align}
  G_{\text{ads}}X &= a_X \cdot G_{\text{ads}}{\text{CO}} + c_X
  && \text{(where  $X$ = COOH, CHO, CH\textsubscript{2}O)}  \label{supp_eq1:scaling_relation_x_co}  \\
  G_{\text{ads}}Y &= a_Y \cdot G_{\text{ads}}{\text{OH}} + c_Y
  && \text{(where  $Y$ = O, OCH\textsubscript{3})}          \label{supp_eq2:scaling_relation_y_oh}
\end{align}

Here, CO and OH serve as descriptors for C-centered ($X$) and O-centered ($Y$) groups, respectively, and $a_{\mathrm{X}}, c_{\mathrm{X}}, a_{\mathrm{Y}}, c_{\mathrm{Y}}$ are species-specific scaling parameters.
The scaling relations simplify the adsorption energy landscape to a two-dimensional space, thereby enabling visualization.

However, it's crucial to note that in the CO\textsubscript{2}RR process, most species are not solely C- or O-centered.
Thus, it becomes advantageous to incorporate both types of species in the scaling relation.
Building upon the hybrid scaling relation introduced in the main text, the free energy change ($\Delta\mathit{G}$) for any given reaction step can be readily calculated using \cref{main_eq14:general_scaling_relation} from the main text.
Consider the generic reaction step:
\begin{align}
  \ce{\text{*}A + H+ + e- -> \text{*}B}  \label{supp_eq3:generic_reaction}
\end{align}

The free energy change $\Delta\mathit{G}$ is defined as:
\begin{align}
  \Delta G = G(\text{*}\mathrm{B}) - G(\text{*}\mathrm{A}) - G(\mathrm{H}^+) - G(e^-)  \label{supp_eq4:free_energy_change}
\end{align}

Focusing on any reaction step delineated by \cref{main_eq1:co2rr1} to \cref{main_eq8:co2rr8} in the main text,
\cref{main_eq15:limiting_potential} shows that limiting potential $U_{\mathrm{L}}$ at a specific external potential $U$ is fully determined by the adsorption energies of two descriptors.
For instance, for the reaction described by \cref{main_eq2:co2rr2}:
\begin{align}
  \Delta G &= G(\text{*COOH}) - G(\text{*}) - G(\text{CO}_2) - G(\text{H}^+) - G(\text{e}^-)  \\
  &= G(\text{COOH}) + G(\text{*}) + G_{\text{ads}}\text{COOH} -
  G(\text{*}) - G(\text{CO}_2) -G(\text{H}^+) - G(\text{e}^-) \\
  &= G_{\text{ads}}\text{COOH} + G(\text{COOH}) - G(\text{CO}_2) -G(\text{H}^+) - G(\text{e}^-)  \label{supp_eq5:exp_reaction}
\end{align}

From adsorption energy linear relations, we have:
\begin{align}
  G_{ads}{\mathrm{COOH}} = a_{\text{COOH}} \cdot G_{ads}{\mathrm{CO}} +
  b_{\text{COOH}} \cdot G_{ads}{\mathrm{OH}} + c_{\text{COOH}}  \label{supp_eq6:eads}
\end{align}

which then reduces Supplementary \cref{supp_eq5:exp_reaction} to:
\begin{align}
  \begin{split}
    \Delta G = &(a_{\text{COOH}} \cdot G_{\text{ads}}{\mathrm{CO}} +
    b_{\text{COOH}} \cdot G_{\text{ads}}{\mathrm{OH}}) + \\
    &[c_{\text{COOH}} + G(\mathrm{COOH}) -
    G(\mathrm{CO}_2) - G(\mathrm{H}^+) - G(\mathrm{e}^-)]  \label{supp_eq7:reduced_eads}
  \end{split}
\end{align}
In this case, the first term depends solely on the descriptors, while the second term remains constant at a given potential $U$.

% Supplementary Table 13: Eads scaling relation params from hybrid descriptor method
\begin{table}[htbp]
\label{supp_table13:scaling_params}
  \caption{Adsorption free energy scaling relation parameters
    determined by the hybrid descriptor method.}
  \center
  \small
  \begin{tabularx}{0.75\textwidth}{@{}l *{3}{X} @{}}
    \toprule
    Adsorbate             & $a_Z$   & $b_Z$   & $c_Z$    \\
    \midrule
    COOH                  & 0.4032  & 0.5132  &  0.0899  \\
    CHO                   & 0.5977  & 0.3664  & -2.6763  \\
    CH\textsubscript{2}O  & 0.5011  & 0.5011  & -0.9806  \\
    OCH\textsubscript{3}  & 0.0532  & 1.0114  &  0.8489  \\
    O                     & 0.3758  & 1.2580  &  0.0910  \\
    H                     & 0.3164  & 0.3568  & -0.6764  \\
    \bottomrule
  \end{tabularx}

  \smallskip

  \begin{flushright}
    \begin{minipage}{\textwidth}
      \footnotesize\textit{Note:} $a_{Z}$, $b_{Z}$ and $c_{Z}$ correspond
      to coefficients for $G_{\text{ads}}\text{CO}$, $G_{\text{ads}}\text{OH}$
      and constant terms, where $Z$ stands for the adsorbed species of interest.
  \end{minipage}
  \end{flushright}
\end{table}

% SI Figure 10: Improvement in R2 from hybrid-descriptor method
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig10_r2_hyb_des.pdf}
  \caption{\textbf{$R^2$ improvement from hybrid-descriptor method.}
  Improvement in the coefficient of determination ($R^2$) with x-axis
  as CO ratio in CO/OH hybrid descriptors.
  (\textbf{a}) averaged $R^2$ and those of (\textbf{b}) CO\textsubscript{2},
  (\textbf{c}) COOH,
  (\textbf{d}) CHO, (\textbf{e}) CH\textsubscript{2}O, (\textbf{f})
  OCH\textsubscript{3}, (\textbf{g}) O and (\textbf{h}) H are shown.}
  \label{supp_fig10:r2_hyb_des}
\end{figure}

% Supplementary Table 14: Performance comparison of hybrid/single descriptors
\begin{table}[htbp]
\label{supp_table14:hybrid_des_perf}
  \caption{Performance comparison of hybrid descriptor and single
    descriptor for linear scaling relations.}
  \small
  \center
  \begin{tabularx}{0.75\textwidth}{@{}l *{3}{X} @{}}
    \toprule
    \multirow{2}{*}{Adsorbate} & Hybrid Descriptor $R^2$ & Single Descriptor $R^2$ & \multirow{2}{*}{$R^2$} \\
                               & (Optimal CO Ratio)      & Reference Species       &                        \\
    \midrule
    CO\textsubscript{2}   & 0.8083 (57\%)  & 0.7368 (CO)  & 0.0715  \\
    COOH                  & 0.9500 (44\%)  & 0.8113 (CO)  & 0.1387  \\
    CHO                   & 0.9146 (62\%)  & 0.8541 (CO)  & 0.0605  \\
    CH\textsubscript{2}O  & 0.9299 (50\%)  & 0.8226 (CO)  & 0.1073  \\
    OCH\textsubscript{3}  & 0.9936 (5\%)   & 0.9929 (OH)  & 0.0007  \\
    O                     & 0.8974 (23\%)  & 0.8825 (OH)  & 0.0149  \\
    H                     & 0.8303 (47\%)  & 0.7617 (OH)  & 0.0686  \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Volcano plots}
\label{supp_sec2.6_volcano}

In addition to the selectivity volcano plot featured in the main text, we provide a comprehensive plot capturing all rate-determining steps for CO\textsubscript{2}RR in \cref{supp_fig11:rds_volcano}. A separate plot focusing on selectivity over HER is also presented in \cref{supp_fig12:sel_volc}.

% SI Figure 11: CO2RR rate determining step volcano plot
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig11_rds_volcano.pdf}
  \caption{\textbf{CO\textsubscript{2}RR rate determining step volcano plot.}
  This graph illustrates the rate-determining step associated with each region within the CO\textsubscript{2}RR process.}
  \label{supp_fig11:rds_volcano}
\end{figure}

% SI Figure 12: CO2RR selectivity volcano plot
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig12_sel_volc.pdf}
  \caption{\textbf{CO\textsubscript{2}RR selectivity volcano plot.}
  Selectivity volcano plot computed from linear scaling relations.
  In the plot, the circle, triangle, square, diamond, plus, and star symbols correspond to
  metals supported on g-C\textsubscript{3}N\textsubscript{4}, nitrogen-doped graphene, graphene with dual-vacancy,
  black phosphorous, BN, and C\textsubscript{2}N, respectively.}
  \label{supp_fig12:sel_volc}
\end{figure}

\subsection{Activity periodic tables}
\label{supp_sec2.7_ptable}

% SI Figure 13: Limiting potential periodic table for nitrogen-doped graphene
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig13_n-gra_ptable.pdf}
  \caption{\textbf{Limiting potential periodic table for nitrogen-doped graphene.}}
  \label{supp_fig13:n-gra_ptable}
\end{figure}

% SI Figure 14: Limiting potential periodic table for graphene with dual-vacancy
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig14_gra-vac_ptable.pdf}
  \caption{\textbf{Limiting potential periodic table for graphene with dual-vacancy.}}
  \label{supp_fig14:gra-vac_ptable}
\end{figure}

% SI Figure 15: Limiting potential periodic table for black phosphorus
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig15_BP_ptable.pdf}
  \caption{\textbf{Limiting potential periodic table for black phosphorus.}}
  \label{supp_fig15:BP_ptable}
\end{figure}

% SI Figure 16: Limiting potential periodic table for boron nitride
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig16_BN_ptable.pdf}
  \caption{\textbf{Limiting potential periodic table for boron nitride.}}
  \label{supp_fig16:BN_ptable}
\end{figure}

% SI Figure 17: Limiting potential periodic table for monolayer C2N
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig17_C2N_ptable.pdf}
  \caption{\textbf{Limiting potential periodic table for monolayer C\textsubscript{2}N.}}
  \label{supp_fig17:C2N_ptable}
\end{figure}

\newpage
% SI Section Two: Additional details on Machine Learning

\section{Additional details on machine learning method}

\subsection{d-band center and adsorption energy relations}
\label{supp_sec3.1_dband_eads}

% SI Figure 18: d-band venter vs adsorption energy scatter plot
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig18_d-band_eads.pdf}
  \caption{\textbf{Scatter plot of d-Band Center and Adsorption Energy Relationships.}
  This figure delineates the relationship between d-band center and adsorption energies for
  various species: (\textbf{a}) CO\textsubscript{2}, (\textbf{b}) COOH, (\textbf{c}) CO,
  (\textbf{d}) CHO, (\textbf{e}) CH\textsubscript{2}O, (\textbf{f}) OCH\textsubscript{3},
  (\textbf{g}) O, and (\textbf{h}) OH.
  Data from all six substrates investigated in this study are included.}
  \label{supp_fig18:dband_vs_eads}
\end{figure}

\subsection{Feature correlation analysis}
\label{supp_sec3.2_feature_corr}

\cref{supp_fig19:pairwise_eads_des} displays the pairwise relationships between CO adsorption energy
and its four most correlated descriptors: d-band center (spin-up) $\delta\epsilon_{\text{d}}\uparrow$,
lattice parameter $\gamma$, vacuum level $E_\text{vac}$, and electronegativity
on the Allen scale $\chi_\text{Allen}$, identified by Kendall rank correlation analysis.
\cref{supp_table15:descriptor_notations} lists notations for elementary descriptors and electronic descriptors.

% SI Figure 19: Pairwise plot of adsorption energy vs key descriptors
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig19_pairwise_eads_des.pdf}
  \caption{\textbf{Pairwise Plot of Adsorption Energy and Key Descriptors.}
  This chart displays the pairwise relationships between CO adsorption energy and
  its four most correlated descriptors: d-band center (spin-up) $\delta\epsilon_{\text{d}}\uparrow$,
  lattice parameter $\gamma$, vacuum level $E_\text{vac}$, and electronegativity
  on the Allen scale $\chi_\text{Allen}$, identified by Kendall rank correlation analysis.
  Data from all six substrates investigated in this study are included.}
  \label{supp_fig19:pairwise_eads_des}
\end{figure}

% Supplementary Table 15: Notations for elementary descriptors and electronic descriptors
\begin{table}[htbp]
\label{supp_table15:descriptor_notations}
  \caption{Notations for elementary descriptors and electronic descriptors.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{llll}
      \toprule
      \multicolumn{2}{c}{Elementary Descriptors}       & \multicolumn{2}{c}{Electronic Descriptors}            \\
      \midrule
      $R$            & atomic radius                   & $\alpha$, $\beta$, $\gamma$  & lattice parameters     \\
      $E_\text{ea}$  & electron affinity               & $\Phi_{\text{DFT}}$    & DFT calculated work function \\
      $\Phi$         & work function                   & $E_\text{vac}$         & vacuum level                 \\
      $\chi_\text{Allen}$, $\chi_\text{P}$, $\chi_\text{RevP}$
      & \parbox[t]{4cm}{electronegativity in Allen, Pauling and revised Pauling scales}
      & $E_\text{g}$, $E_\text{g}\uparrow$, $E_\text{g}\uparrow$
      & \parbox[t]{4cm}{average, spin-up and spin-down band gaps}                                              \\
      $A_\text{r}$   & relative atomic mass            & $e_\text{d}$           & number of d-electrons        \\
      $E_\text{i}$   & ionization energy & $\delta\epsilon_{\text{d}}\uparrow$  & d-band centre (spin-up)      \\
      $G$            & group number                    & $W_\text{d}$           & d-band width                 \\
      $P$            & period number                   & $e_{\text{Bader}}$     & Bader charge                 \\
      $V$            & number of valence electrons     &                        &                              \\
      \bottomrule
    \end{tabular}
  }
\end{table}

\subsection{Input pipeline construction}
\label{supp_sec3.3_input_pipeline}

The format of the 2D array for the eDOS is naturally well-suited for use as neural network inputs.
In this study, the array adopts a shape of [numSamplings, numOrbitals, numChannels], as illustrated in \cref{supp_fig20:input_dos}.
Here, ``numSamplings'' represents the number of data points sampled across the energy spectrum,
``numOrbitals'' denotes the number of DOS orbitals,
and ``numChannels'' indicates the number of atomic channels.

% SI Figure 20: Representation of the input eDOS data structure
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.3\textwidth]{supp_fig20_input_dos.pdf}
  \caption{\textbf{Schematic representation of the input eDOS data structure.}}
  \label{supp_fig20:input_dos}
\end{figure}

For practical purposes, certain spatial hierarchies are expected to remain invariant,
such as the energy range over which the eDOS is sampled.
In this work, we maintain a consistent energy range of -14 eV to 6 eV,
sampled at a density of 50 \si{\text{points} \cdot \text{eV}^{-1}}.
In addition to maintaining a fixed energy range,
the following steps are taken to standardize the input eDOS data:
  1.	For atoms lacking d-electrons, the eDOS is zero-padded.
  2.	f-orbital eDOS data are excluded.
  3.	The eDOS of adsorbates is zero-padded along the Channels axis to align with the adsorbate that contains the maximum number of atoms, which, in this study, is the *OCH\textsubscript{3} species. This padded eDOS is then appended to the metal eDOS.

As a result of these preprocessing steps, the input eDOS arrays consistently have a shape of [4000, 9, 6].
This corresponds to 4000 sampling points across the energy range, 9 eDOS orbitals, and 6 atomic channels.

\subsection{Hyperparameter tuning for CNN}
\label{supp_sec3.4_hyperparam}

% Supplementary Table 16: The hyperparameter search space
\begin{table}[htbp]
\label{supp_table16:hyperparam_space}
  \caption{The hyperparameter search space.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{lll}
      \toprule
      Hyperparameter& Search Space              & Description                                                       \\
      \midrule
      learningRate  & [10\textsuperscript{-4}, 10\textsuperscript{-2}], log scale & Learning rate of Adam \cite{kingma2017adam} optimizer \\
      dropoutRate   & [0, 1]                    & Dropout rate                                                      \\
      root\_fc0     & 128, 256, 512             & Output dimensionality of 1\textsuperscript{st} fully connected layer in root           \\
      root\_fc1     & 256, 512, 1024, 2048      & Output dimensionality of 2\textsuperscript{nd} fully connected layer in root           \\
      root\_act     & ``tanh'', ``relu'', ``sigmoid'' & Activation function of fully connected layers in root             \\
      kernelSize    & [2, 32]                   & Width of the convolution window                                   \\
      numFilters    & [8, 64]                   & Number of filters in the convolution                              \\
      numConvLayers & [1, 16]                   & Number of convolution layers                                      \\
      numConvBlocks & [1, 16]                   & Number of convolution blocks                                      \\
      br\_fc0       & 16, 32, 64, 128, 256, 512 & Output dimensionality of 1\textsuperscript{st} fully connected layer in branch      \\
      br\_fc1       & 8, 16, 32, 64, 128, 256   & Output dimensionality of 2\textsuperscript{nd} fully connected layer in branch      \\
      br\_act       & ``tanh'', ``relu'', ``sigmoid'' & Activation function of fully connected layers in branch           \\
      \bottomrule
    \end{tabular}
  }
\end{table}

% Supplementary Table 17: Optimal hyperparameters identified with Hyperband algorithm
\begin{table}[htbp]
\label{supp_table17:opt_hyperparam}
  \caption{Optimal hyperparameters identified with Hyperband algorithm.}
  \resizebox{\textwidth}{!}{%
    \begin{tabular}{lll}
      \toprule
      Hyperparameter& Optimal Value & Description                                              \\
      \midrule
      learningRate  & $10^{-3}$ & Learning rate of Adam \cite{kingma2017adam} optimizer                              \\
      dropoutRate   & 0.3       & Dropout rate                                                 \\
      root\_fc0     & 256       & Output dimensionality of 1\textsuperscript{st} fully connected layer in root   \\
      root\_fc1     & 512       & Output dimensionality of 2\textsuperscript{nd} fully connected layer in root   \\
      root\_act     & ``relu''    & Activation function of fully connected layers in root        \\
      kernelSize    & 10        & Width of convolution window                                  \\
      numFilters    & 17        & Number of filters in the convolution                         \\
      numConvLayers & 3         & Number of convolution layers                                 \\
      numConvBlocks & 1         & Number of convolution blocks                                 \\
      br\_fc0       & 256       & Output dimensionality of 1\textsuperscript{st} fully connected layer in branch \\
      br\_fc1       & 128       & Output dimensionality of 2\textsuperscript{nd} fully connected layer in branch \\
      br\_act       & ``tanh''    & Activation function of fully connected layers in branch      \\
      \bottomrule
    \end{tabular}
  }
\end{table}

% Supplementary Table 18: Prediction MAEs of the CNN model
\begin{table}[htbp]
\label{supp_table18:cnn_mae}
  \caption{Prediction mean absolute errors of the CNN model.}
  \small
  \center
  \begin{tabularx}{0.75\textwidth}{@{}lXX@{}}
    \toprule
    Adsorbate             & Original dataset (eV)  & Augmented dataset (eV)  \\
    \midrule
    CO\textsubscript{2}   & 0.0447                 & 0.0336                  \\
    COOH                  & 0.0436                 & 0.0581                  \\
    CO                    & 0.0402                 & 0.0402                  \\
    CHO                   & 0.0614                 & 0.0778                  \\
    CH\textsubscript{2}O  & 0.0553                 & 0.0632                  \\
    OCH\textsubscript{3}  & 0.0605                 & 0.0636                  \\
    O                     & 0.1135                 & 0.1081                  \\
    OH                    & 0.0547                 & 0.0602                  \\
    H                     & 0.0703                 & 0.0646                  \\
    \bottomrule
  \end{tabularx}

  \smallskip

  \begin{flushright}
  \begin{minipage}{\textwidth}
    \footnotesize\textit{Note:} The same CNN model, trained on the
      augmented dataset, was used for all predictions. The term ``original dataset''
      refers to evaluations conducted on the original dataset without any augmented data;
      it does not imply the use of a different CNN model trained solely on the original dataset.
      Similarly, ``augmented dataset'' refers to evaluations on the dataset with
      both original data and augment data, not a distinct model.
  \end{minipage}
  \end{flushright}
\end{table}

\subsection{Occlusion experiment}
\label{supp_sec3.5_occlusion}

% SI Figure 21: Illustration of the eDOS occlusion experiment
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig21_occlusion.pdf}
  \caption{\textbf{Illustration of the eDOS occlusion experiment.}
  This figure demonstrates the step-by-step process of the occlusion experiment.
  The input eDOS array first undergoes zero-padded, and then a masker of width ``width'' is applied,
  resulting in an occlusion array that preserves the shape of the input array.}
  \label{supp_fig21:occlusion}
\end{figure}

% SI Figure 22: Occlusion experiment with a masker width of 1
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{supp_fig22_occl_wid1.pdf}
  \caption{\textbf{Occlusion Experiment with a masker width of ``1''.}
  This figure shows an occlusion experiment on the Ge atom in
  the initial state of CO adsorption on Ge@g-C\textsubscript{3}N\textsubscript{4}.}
  \label{supp_fig22:occl_wid1}
\end{figure}

% SI Figure 23: Occlusion experiment with a masker width of 3
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{supp_fig23_occl_wid3.pdf}
  \caption{\textbf{Occlusion Experiment with a masker width of ``3''.}
  This figure shows an occlusion experiment on the Ge atom in
  the initial state of CO adsorption on Ge@g-C\textsubscript{3}N\textsubscript{4}.}
  \label{supp_fig23:occl_wid3}
\end{figure}

% SI Figure 24: Occlusion experiment with a masker width of 5
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{supp_fig24_occl_wid5.pdf}
  \caption{\textbf{Occlusion Experiment with a masker width of ``5''.}
  This figure shows an occlusion experiment on the Ge atom in
  the initial state of CO adsorption on Ge@g-C\textsubscript{3}N\textsubscript{4}.}
  \label{supp_fig24:occl_wid5}
\end{figure}

% SI Figure 25: Occlusion experiment with a masker width of 11
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{supp_fig25_occl_wid11.pdf}
  \caption{\textbf{Occlusion Experiment with a masker width of ``11''.}
  This figure shows an occlusion experiment on the Ge atom in
  the initial state of CO adsorption on Ge@g-C\textsubscript{3}N\textsubscript{4}.}
  \label{supp_fig25:occl_wid11}
\end{figure}

% SI Figure 26: Occlusion experiment with a masker width of 21
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{supp_fig26_occl_wid21.pdf}
  \caption{\textbf{Occlusion Experiment with a masker width of ``21''.}
  This figure shows an occlusion experiment on the Ge atom in
  the initial state of CO adsorption on Ge@g-C\textsubscript{3}N\textsubscript{4}.}
  \label{supp_fig26:occl_wid21}
\end{figure}

% SI Figure 27: Occlusion experiment with a masker width of 31
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{supp_fig27_occl_wid31.pdf}
  \caption{\textbf{Occlusion Experiment with a masker width of ``31''.}
  This figure shows an occlusion experiment on the Ge atom in
  the initial state of CO adsorption on Ge@g-C\textsubscript{3}N\textsubscript{4}.}
  \label{supp_fig27:occl_wid31}
\end{figure}

% SI Figure 28: Occlusion experiment with a masker width of 41
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{supp_fig28_occl_wid41.pdf}
  \caption{\textbf{Occlusion Experiment with a masker width of ``41''.}
  This figure shows an occlusion experiment on the Ge atom in
  the initial state of CO adsorption on Ge@g-C\textsubscript{3}N\textsubscript{4}.}
  \label{supp_fig28:occl_wid41}
\end{figure}

% SI Figure 29: Occlusion experiment with a masker width of 51
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{supp_fig29_occl_wid51.pdf}
  \caption{\textbf{Occlusion Experiment with a masker width of ``51''.}
  This figure shows an occlusion experiment on the Ge atom in
  the initial state of CO adsorption on Ge@g-C\textsubscript{3}N\textsubscript{4}.}
  \label{supp_fig29:occl_wid51}
\end{figure}

\subsection{Shifting experiment}
\label{supp_sec3.6_shifting}

% SI Figure 30: Illustration of the eDOS shifting experiment
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{supp_fig30_shifting.pdf}
  \caption{\textbf{Illustration of the eDOS shifting experiment.}
  This figure illustrates the sequential stages of the shifting experiment.
  Initially, the input eDOS array is shifted along the Energy axis.
  Subsequently, it undergoes zero-padding and cropping,
  ensuring a consistent shape based on the shifting direction, to match the input eDOS array.
  The resulting shifted array is then processed by the CNN model to
  assess the disturbance in adsorption energy caused by the shifting operation.}
  \label{supp_fig30:shifting}
\end{figure}

% SI Figure 31: Shifting experiment on p orbital
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.75\textwidth]{supp_fig31_p_shifting.pdf}
  \caption{\textbf{Shifting experiment on p orbital.}
  Effect of orbital shifting on CO adsorption energy, predicted by the CNN model,
  for single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.
  The perturbations caused by shifting of entire p orbital are presented.
  The shifting step size corresponds to the energy resolution of the eDOS, set at 0.005 eV in this study.
  A positive eDOS shift indicates a shift towards higher energy levels, and vice versa.}
  \label{supp_fig31:p_shifting}
\end{figure}

\newpage
% SI Section Three: Other Additional Information

\section{Other additional information}
\label{supp_sec4_env}

\subsection{Additional information on machine learning environment}
\label{supp_sec4.1_ml_env}

The hyperparameter optimization for the proposed CNN model was performed using NVIDIA\textsuperscript{\textregistered} V100 Tensor Core GPUs,
courtesy of the Australian National Computational Infrastructure's (NCI) HPC Gadi system.
The HyperBand algorithm \cite{li2018hyperband}, as implemented in
KerasTuner \cite{omalley2019kerastuner}, facilitated the tuning process.

% Supplementary Table 19: Packages in the DL env
\begin{table}[htbp]
\label{supp_table19:pack_dl_env}
  \caption{Core packages in the deep learning environment.}
  \small
  \center
  \begin{tabularx}{0.9\textwidth}{@{}l *{3}{X} @{}}
    \toprule
    Name                 & Version  & Build                         & Channel      \\
    \midrule
    conda                & 4.13.0   & py39h06a4308\textunderscore0  & anaconda     \\
    cuda-nvcc            & 11.8.89  & 0                             & nvidia       \\
    cuda-toolkit         & 11.8.0   & 0                             & nvidia       \\
    cudatoolkit          & 11.6.0   & habf752d\textunderscore9      & nvidia       \\
    ipykernel            & 6.9.1    & py39h06a4308\textunderscore0  & anaconda     \\
    ipython              & 7.34.0   & pypi\textunderscore0          & pypi         \\
    jupyter\_client      & 7.2.2    & py39h06a4308\textunderscore0  & anaconda     \\
    jupyterlab           & 3.4.4    & py39h06a4308\textunderscore0  & anaconda     \\
    keras                & 2.9.0    & pypi\textunderscore0          & pypi         \\
    keras-preprocessing  & 1.1.2    & pypi\textunderscore0          & pypi         \\
    keras-tuner          & 1.3.5    &                               & pypi         \\
    matplotlib           & 3.5.1    & py39h06a4308\textunderscore1  & anaconda     \\
    numpy                & 1.22.4   & pypi\textunderscore0          & pypi         \\
    pandas               & 1.3.4    & py39h8c16a72\textunderscore0  & anaconda     \\
    python               & 3.9.12   & h12debd9\textunderscore1      & anaconda     \\
    scikit-learn         & 1.1.1    & py39h6a678d5\textunderscore0  & anaconda     \\
    scipy                & 1.8.1    & py39hddc5342\textunderscore3  & conda-forge  \\
    tensorboard          & 2.9.1    & pypi\textunderscore0          & pypi         \\
    tensorflow           & 2.9.3    & pypi\textunderscore0          & pypi         \\
    tensorflow-gpu       & 2.9.3    & pypi\textunderscore0          & pypi         \\
    xlrd                 & 2.0.1    & pyhd3eb1b0\textunderscore0    & anaconda     \\
    yaml                 & 0.2.5    & h7b6447c\textunderscore0      & anaconda     \\
    \bottomrule
  \end{tabularx}

  \smallskip

  \begin{flushright}
  \begin{minipage}{\textwidth}
    \footnotesize\textit{Note:} The complete list of packages is
      provided along with the source code.
  \end{minipage}
  \end{flushright}
\end{table}

\subsection{Additional information on data analysis and visualization environment}
\label{supp_sec4.2_vis_env}

Data analysis and visualization tasks were executed on an Apple MacBook Air featuring an octa-core M2 chip.
For accuracy, the TensorFlow-Metal plugin for Mac-based GPU acceleration was deliberately excluded,
as it could potentially yield erroneous CNN model predictions.

The illustrations of CNN architecture displayed in \cref{main_fig1:pipeline} of the main text,
as well as \cref{supp_fig21:occlusion} and \cref{supp_fig30:shifting},
were obtained from the GitHub repository dair-ai/ml-visuals \cite{Saravia_ML_Visuals_2021}.
Blender was used to create and render the visual representation of the
catalyst analysis pipeline shown in \cref{main_fig1:pipeline} of the main text.

% Supplementary Table 20: Packages in data analysis and vis env
\begin{table}[htbp]
\label{supp_table20:pack_vis_env}
  \caption{Core packages in data analysis and visualization environment.}
  \small
  \center
  \begin{tabularx}{0.9\textwidth}{@{}l *{3}{X} @{}}
    \toprule
    Name              & Version  & Build                     & Channel  \\
    \midrule
    bokeh             & 3.2.2    & pypi\textunderscore0      & pypi     \\
    colorcet          & 3.0.1    & pypi\textunderscore0      & pypi     \\
    heatmapz          & 0.0.4    & pypi\textunderscore0      & pypi     \\
    keras             & 2.13.1   & pypi\textunderscore0      & pypi     \\
    keras-tuner       & 1.3.5    & pypi\textunderscore0      & pypi     \\
    matplotlib        & 3.7.2    & pypi\textunderscore0      & pypi     \\
    numpy             & 1.24.3   & pypi\textunderscore0      & pypi     \\
    pandas            & 2.0.3    & pypi\textunderscore0      & pypi     \\
    python            & 3.11.4   & hb885b13\textunderscore0  &          \\
    pyyaml            & 6.0.1    & pypi\textunderscore0      & pypi     \\
    scikit-learn      & 1.3.0    & pypi\textunderscore0      & pypi     \\
    scipy             & 1.11.1   & pypi\textunderscore0      & pypi     \\
    seaborn           & 0.12.2   & pypi\textunderscore0      & pypi     \\
    tensorboard       & 2.13.0   & pypi\textunderscore0      & pypi     \\
    tensorflow        & 2.13.0   & pypi\textunderscore0      & pypi     \\
    tensorflow-macos  & 2.13.0   & pypi\textunderscore0      & pypi     \\
    \bottomrule
  \end{tabularx}

  \smallskip

  \begin{flushright}
  \begin{minipage}{\textwidth}
    \footnotesize\textit{Note:} The complete list of packages
      is provided along with the source code.
  \end{minipage}
  \end{flushright}
\end{table}

% End of Supplementary Information